---
title: SIER models for rabies epidemic in China
author: Martina Iannacito\textsuperscript{1} and Matteo Ciciani\textsuperscript{2}
affiliation:
  - \textsuperscript{1}\textit{...}
  - \textsuperscript{2}\textit{Master in Quantitative and Computational Biology, University of Trento, Italy}
output:
    pdf_document:
      latex_engine: xelatex
      template: SelfArx-template.tex
abstract: Rabies epidemic in china. reporduce some of the results presented in \cite{Cina}. Novel results.
---
# Introduction

## The biology of rabies
Rabies is one of the oldest diseases in recorded human history, dating back more
than 4000 years. More than 70000 human rabies deaths occur annually, and Asia
accounts for approximately 80\% of the worldwide total \cite{Wang}. As 40\% of
those infected are children, the years of life lost by this infection makes
rabies the seventh most important global infectious disease \cite{Schnell}.
Currently, China is second only to India in the number of people killed by
rabies worldwide, and the epidemic area has expanded to almost the entire
country \cite{Wang}.

Rabies is an acute encephalomyelitis (inflammation of the brain and the spinal
cord) caused by viruses in the genus \emph{Lyssavirus} of the family
\emph{Rhabdoviridae}. Rabies lyssavirus (RABLV), the prototype species of the
genus, causes disease in nearly all terrestrial mammals. Despite the worldwide
distribution of RABLV, only a few animal species in the orders \emph{Carnivora},
\emph{Chiroptera} (bats) and \emph{Primates} play a critical role in maintaining
the circulation of the virus in nature. Most of these mammals harbor
species-specific RABLV variants which circulate within discrete geographic
boundaries. However, because almost all cases of human rabies result from the
bites of rabid dogs, dog-maintained viruses are by far the most important target
of control measures \cite{Velasco}.

As for most viruses, the rabies virus life cycle can be divided into three
phases. The first phase includes entry into the host cell by endocytosis
followed by release of the viral genome into the cytoplasm. The classical
wild-type rabies virus first infects a muscle cell, then enters a motor neuron
at the neuromuscular junction. Rabies virus particles are subsequently
transported in a retrograde direction through the axon of the infected neuron.
Once the virus particles reach the cell body of the neuron, the second phase
commences, which involves the production of the virion components. The last
phase of the life cycle comprises the assembly of the viral components and the
release of mature rabies virus particles, which can then start a new round of
infection \cite{Schnell}.

By the time a patient develops the first prodromal symptoms, such as fever,
flu-like symptoms, and gastrointestinal disturbances, the virus is already
widely disseminated throughout the central nervous system \cite{Hemachudha}.

Clinical manifestation of rabies in humans has two forms: the furious form (80\%
of infections), and the paralytic form (20\% of infections). The furious form of
rabies is characterized by hydrophobia, episodes of hallucinations and
excitement. These behavioral changes occur simultaneously with the shedding of
large amounts of rabies virus in the saliva, which facilitates the spread of the
virus to a new host. The paralytic form of rabies is characterized by weakness
and flaccid paralysis. In both cases, survival after the onset of symptoms is
rarely more than 7 days and the disease is lethal in almost 100\% of cases
\cite{Schnell}.

Human rabies virus infections are controlled by limiting the infection in
animals and by post-exposure prophylaxis (PEP) of humans after contact with a
potentially infected animal. Current PEP for previously non-immunized
individuals consists of one dose with rabies immunoglobulin in conjunction with
five doses of immunization with inactivated rabies virus vaccines. This regimen
is effective, but its usefulness in developing countries, where rabies is
endemic, is hampered by the high cost and lack of compliance \cite{Schnell}.

## SEIR models

One of the most known model used to investigate the evolution of epidemics is
SEIR. It is coherent to apply a basic SEIR model if the disease considered has
an average duration shorter than the average time needed to observe significant
modifications in the population structure. For example it can be applied to
describe flu epidemics in the human species, since the average lifespan of a
human being and the time needed to generate a new individual are on average
longer than the duration of the flu. Under this assumption demographical aspects
are negligible and consequently the model focuses only on the disease dynamic.
In order to apply the model the total population $N(t)$ at time t is split into
four classes:
\begin{itemize}
  \item $S(t)$ : the class of individuals susceptible to the disease at time t;
  \item $E(t)$ : the class of individuals exposed to the pathogen and not
  contagious at time t;
  \item $I(t)$ : the class of infectious individuals at time t;
  \item $R(t)$ : the class of removed individuals, i.e. those who get out of the
  model because they are dead or they have become immune at time t.
\end{itemize}

Then it is assumed that susceptible individuals who come into contact with
infected ones become exposed at rate $\lambda(t)$ , exposed move into the
infectious class at rate $k(t)$ and at last infectious are removed at rate
$\gamma(t)$ . The force of infection $\lambda(t)$ is mathematically described
as:

$$
\lambda(t) = c(t)\chi\frac{I(t)}{N(t)}
$$

where $c(t)$ is the average number of contact of one individual in the unit of
time, $\chi$ is the probability to become infect after a contact with an
infectious and $\frac{I(t)}{N(t)}$ is the total number of infectious. To make
computations easier it is assumed that all these parameter are constant in time
and thanks to the absence of demographic aspects the total population is also
constant. Therefore the SEIR ODE system is:

\begin{align}
  \label{s1}
  \begin{cases}
    S'(t) = -\frac{c\chi}{N}S(t)I(t)\\
    E'(t) = \frac{c\chi}{N}S(t)I(t)-kE(t)\\
    I'(t) = kE(t)-\gamma I(t)\\
    R'(t) =\gamma I(t)
  \end{cases}
\end{align}

It is important to underline that if $\gamma$ is the rate at which an infectious
is removed, then $\gamma$ is the average time an individual spends as
infectious. Consequently the basic epidemiological reproduction number is the
number of ill individuals generated by an infectious in the life time as
infectious, i.e.

$$
R_0 = \frac{c\chi}{\gamma}
$$

It is possible to obtain a simpler system rescaling \ref{s1} in the following
way: $\tau = \gamma t$, $u(t) =\frac{S(t)}{N}$, $v(t) = \frac{E(t)}{N}$, $w(t) =
\frac{I(t)}{N}$ and $z(t) =\frac{R(t)}{N}$. Thanks to the applied rescaling,
it follows that $u+v+w+z=\frac{S+E+I+R}{N}=1$. Denoting
$\dot{u}=\frac{du}{d\tau}$, the system (\ref{s1}) becomes:
\begin{align*}
	\begin{cases}
		\dot{u}(\tau) =-\frac{c\chi}{\gamma}u(\tau)w(\tau)\\
		\dot{v}(\tau) = \frac{c\chi}{\gamma}u(\tau)w(\tau)-\rho v(\tau)\\
		\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
		\dot{z}(\tau) = w(\tau)
	\end{cases}
\end{align*}

which is equivalent to
\begin{align}
	\label{s2}
	\begin{cases}
		\dot{u}(\tau) = -R_0u(\tau)u(\tau)w(\tau)\\
		\dot{v}(\tau) = R_0u(\tau)w(\tau)-\rho v(\tau)\\
		\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
		\dot{z}(\tau) = w(\tau).
	\end{cases}
\end{align}

It can be proved as in \cite{mate1} that given positive initial condition for
the system (\ref{s2}) there is a unique positive solution.

In order to make the model more realistic, it is convenient to introduce a
simple vital dynamic with death rate $\mu$ equal to birth rate $\beta$. Under
this hypothesis the population is still constant in time. The resulting system
is:
\begin{align}
  \label{s3}
	\begin{cases}
    \dot{u}(\tau) = -R_0u(\tau)u(\tau)w(\tau)+\beta(1-u(\tau))\\
    \dot{v}(\tau) = R_0u(\tau)w(\tau)-(\rho+\beta)v(\tau)\\
    \dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
    \dot{z}(\tau) = w(\tau)-\beta z.
  \end{cases}
\end{align}

Reminding that $u+v+z+w=1$, it is easy to compute the equilibria of the system
(\ref{s3}), which are:
\begin{itemize}
	\item the disease free equilibrium $E_0=(1,0,0,0)$;
	\item the endemic equilibrium $E^*=(u^*,v^*,w^*,z^*)$, that exists if and only
	if holds the condition
	\begin{equation}
  	\label{c1}
  	R_0>1+\frac{\beta}{\rho}
	\end{equation}
\end{itemize}

To investigate the stability of the equilibria, it is useful to consider the
jacobian of the reduced system:
\begin{align*}
	\begin{cases}
  	\dot{u}(\tau)=-R_0u(\tau)u(\tau)w(\tau)+\beta(1-u(\tau))\\
  	\dot{v}(\tau) = R_0u(\tau)w(\tau)-(\rho+\beta)v(\tau)\\
  	\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
	\end{cases}
\end{align*}

For the disease free equilibrium the Jacobian is
$$
J(E_0)=\begin{bmatrix} 
-\beta & 0  & -R_0 \\ 
0 &	-(\rho+\beta)	& R_0\\
0 &	\rho	& -1\\
\end{bmatrix}
$$

It can be observed that it is a block matrix. Consequently an eigenvalue is
$\lambda_1=-\beta$, the other ones are the eigenvalues of the matrix
$$
J(E_0)=\begin{bmatrix} 
-(\rho+\beta)	& R_0\\
\rho	& -1\\
\end{bmatrix}
$$
which has negative trace and positive determinant if the condition (\ref{c1})
is false. For the endemic equilibrium the Jacobian is
$$
J(E^*)=\begin{bmatrix} 
-(R_0w^*+\beta) & 0  & -R_0u^* \\ 
R_0w^* &	-(\rho+\beta)	& R_0u^*\\
0 &	\rho	& -1\\
\end{bmatrix}
$$
Using the Routh-Hurwitz conditions, it can be proved as in \cite{mate1} that the
endemic equilibrium, when it exists, is asymptotically stable.

Numerical simulation for the a generic SEIR model are shown in Fig.
\ref{fig:gen1} and \ref{fig:gen2}, for values of $R_0>1$ and $R_0<1$,
respectively.

```{r, echo=FALSE, fig.width=7, fig.height=6,fig.cap="\\label{fig:gen1}Numerical simulation of a SEIR model with $R_0>1$. We can see that the system convergest to the endemic equilibrium."}
source("SEIR_general.R")

ggplot(data = SEIR_out_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab("E, I, R") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*2, name = "S")) + 
  theme(legend.position="bottom", axis.title.y.right =
      element_text(margin = margin(l = 20))) + 
  xlab("time")
```

```{r, echo=FALSE, fig.width=7, fig.height=6,fig.cap="\\label{fig:gen2}Numerical simulation of a SEIR model with $R_0<1$. We can see that the system convergest to the disease-free equilibrium."}
source("SEIR_general.R")

ggplot(data = SEIR_out_long2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab("E, I, R") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*10, name = "S")) + 
  theme(legend.position="bottom", axis.title.y.right =
      element_text(margin = margin(l = 20))) + 
  xlab("time")
```

The basic structure of the SEIR system can be adequately modified and used to
describe the rabies epidemic behavior as in \cite{PAPERNOSTRO?}.

# Reproduction of original results

## SEIR model for dog-human interactions

As it has been pointed out in the biological introduction to rabies, following a
bite from an infectious individual, the virus has an incubation period in the
new host. Consequently the bitten host can be considered exposed. It is
therefore appropriate to describe this epidemic phenomena through a SEIR model
for the dogs and the human population. The most important difference with the
classical SEIR is that in this model only the exposed individuals that are
subjected to PEP can move into the removed class. Moreover, the removed can
loose immunity and get back to susceptible class. The ODE system is then:

\begingroup
\setlength{\mathindent}{0pt}
\begin{align*}
	\begin{cases}
		S_d'(t) = A + \lambda_dR_d + \sigma_d (1 - \gamma_d )E_d - \beta_d S_dI_d - (m_d + k_d)S_d\\
		E_d'(t) = \beta_d S_dI_d - \sigma_d (1 - \gamma_d )E_d - \sigma_d\gamma_d E_d - (m_d + k_d)E_d\\
		I_d'(t) = \sigma_d\gamma_d E_d - (m_d + \mu_d)I_d\\
		R_d'(t) = k(S_d + E_d) - (m_d + \lambda_d)R_d\\
		S_h'(t) = B + \lambda_hR_h + \sigma_h(1 - \gamma_h )E_h - m_hS_h - \beta_{dh}S_hI_h\\
		E_h'(t) = \beta_{dh}S_hI_d - \sigma_h(1 - \gamma_h )E_h - \sigma_h\gamma_hE_h - (m_h + k_h )E_h\\
		I_h'(t) = \sigma_h\gamma_hE_h - (m_h + \mu_h)I_h\\\
		R_h'(t) = k_hE_h - (m_h + \lambda_h )R_h
	\end{cases}
\end{align*}
\endgroup

The flow chart of the model and the respective parameters value are reported
in \cite{Cina}. The system presents both the vitality and the epidemic
dynamic. It is important to underline that the rate of loss of immunity for the
dogs population has been added, compared to the original SEIR. The basic
epidemic reproduction number can be computed \cite{23,96PapCinaBio} as:

$$
R_0=\frac{\beta_d\sigma_d\gamma_d}{(m_d+\kappa_d+\sigma_d)(m_d+\mu_d)}\frac{(m_d
+\lambda_d)A}{(m_d+\kappa_d+\lambda_d)m_d}
$$

Moreover, it can be proved that there are two equilibria:
\begin{itemize}
	\item a disease free equilibrium $E_0=(S_{d,0},0,0,R_{d,0},S_{h,0},0,0,0)$;
	\item an endemic equilibrium
	$E^*=(S^*_d,E^*_d,I^*_d,R^*_{d},S^*_{h},E^*_h,I^*_h,R^*_h)$ which exists if 
	and only if holds the condition
	\begin{equation}
		\label{c2}
		R_0>1.
	\end{equation}
\end{itemize}

The complete expression of $E^*$ as a function of the parameters can be found in
\cite{Cina}. If the condition \ref{c2} is true, it can be proved that $E_0$
is locally asymptotically stable in the region

\begin{align*}
  \Gamma=&\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in(\mathbb{R}_+\cup\{0\})^8~\vert\\
  &0 < S_d + E_d + I_d + R_d \le \frac{A}{m_d}\Bigr\}\\
\end{align*}

Moreover, if \ref{c2} holds, it can be shown that $E*$ is locally asymptotically
stable in $\hat{\Gamma}=\Gamma-\Sigma$, where

$$
\Sigma=\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in\Gamma~\vert~I_h=0\Bigr\}
$$
and all the solution in $\Sigma$ tend toward the disease free equilibrium. A
complete demonstration of the previous statement can be found in
\cite{113papCinaBio}.

Finally, the parameters of the model can be estimated using the data collected
by the Chinese Ministry of Health. Using these parameters, it can be shown that
the $R_0$ value for rabies in China is around 2. Therefore the epidemic moves
towards the endemic equilibrium.

A numerical simulation of the model from 1996, for a period of 50 years, is
shown in Fig. \ref{fig:dh1}. The model indicates the tendency of the rabies
epidemic to persist and move towards the endemic equilibrium.

```{r, echo=FALSE, fig.width=7, fig.height=6,fig.cap="\\label{fig:dh1}Numerical simulations of rabies cases in China from 1996, for a period of 50 years. $S(t)$ and $R(t)$ are plotted in log$_{10}$ scale. The curve for $I(t)$ reproduces Fig. 2.2b of \\cite{Cina}."}
library(latex2exp)
source("SEIR_dogs_humans.R")

# ggplot(data = as.data.frame(SEIR_out)) +
#   geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
#   ylab(TeX("$I_h$")) + xlab("time (months)")

ggplot(data = SEIR_out_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab(latex2exp::TeX("$E_h$, $I_h$")) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~., name =
    latex2exp::TeX("$S_h$, $R_h$"),
    breaks = breaks, labels = labels)) + theme(legend.position="bottom",
      axis.title.y.right = element_text(margin = margin(l = 20))) + 
  xlab("time (months)")
```

We can examine how the initial conditions influence the dynamic of $I(t)$. The
initial conditions of dogs can influence both the number of human rabies cases
and the time of rabies case peak, as shown in Fig. \ref{fig:dh2} Eventually, the
differences become small. We also observe that the peak of the initial outbreak
would be postponed if $S_d(0)$ is decreasing.
 
```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dh2}Number of human rabies cases $I_h(t)$ for different initial values of the susceptible dog population $S_d(0)$. This plot reproduces Fig. 2.3a of \\cite{Cina}."}
ggplot(data = I_h_df_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = latex2exp::TeX("$S_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)") +
  xlim(0, 1000)
```

Instead, the initial conditions of humans influence only the number of human
rabies cases and have no effect on th time of rabies cases peak, as we can see
in Fig. \ref{fig:dh3}.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dh3}Number of human rabies cases $I_h(t)$ for different initial values of the susceptible human population $S_h(0)$. This plot reproduces Fig. 2.3b of \\cite{Cina}."}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = latex2exp::TeX("$S_h(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)") +
  xlim(0, 1000)
```

## SEIV model with domestic and stray dogs

Although the previous model provided a good estimation of the epidemic dynamic
of the last years, it is not completely satisfactory. As a matter of fact,
almost every region of China is populated by stray dogs. Consequently, it is
useful to introduce another model, where the parameters of stray and domestic
dogs are distinctly estimated.

The simplest formulation of this stray-domestic-human system is realized through
a SEIV model with demographic dynamic at constant coefficients. In the SEIV
model the removed class is replaced by the vaccinated class. In this case it is
assumed that only susceptible and exposed domestic flow with a specific rate in
the vaccinated class. Moreover it is assumed that domestic dogs can be abandoned
and consequently the become stray in the respective class. It is also assumed
that only domestic dogs reproduce and that the stray dogs of all classes are
culled at the same rate. For the human population it is coherently supposed that
only exposed individuals are treated, so they can move in the vaccinated class.
The model can be presented as:

\begingroup
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s4}
	\begin{cases}
		S_0'(t) = lS_1 - (\mu + c)S_0 - \beta_0S_0I_0 + p_0E_0 + \delta_1V_0\\
		E_0'(t) = lE_1 + \beta_0S_0I_0 - (\mu + c + \sigma_0 + p_0)E_0\\
		I_0'(t) = \sigma_0E_0 + \epsilon I_1 - (\mu + c + \alpha)I_0\\
		V_0'(t) = lV_1 - (\mu + c + \delta_1)V_0\\
		S_1'(t) = A - \beta S_1I_0 - \beta_1S_1I_1 - (1 + \nu + l)S_1 + p_1E_1 + \delta2V_1\\
		E_1'(t) = \beta S_1I_0 + \beta_1S_1I_1 - (1 + \nu + \sigma_1 + l + p_1)E_1\\
		I_1'(t) = \sigma_1E_1 - (1 + \epsilon + k)I_1\\
		V_1'(t) = \nu(S_1 + E_1) - (1 + l + \delta_2)V_1\\
		S_h'(t) = H - \mu_hS_h - \lambda_{1h}S_hI_0 - \lambda_{2h}S_hI_1 + \delta_{1h}Eh + \delta_{2h}V_h\\
		E_h'(t) =\lambda_{1h}S_hI_0 + \lambda_{2h}S_hI_1 - (\mu_h + \sigma_h + \delta_{1h} + \nu_h)E_h\\
		I_h'(t) = \sigma_hE_h - (\mu_h + \alpha_h )I_h\\
		V_h'(t) = \nu_hE_h - (\mu_h + \delta_{2h})V_h\\
	\end{cases}
\end{align}
\endgroup
	
It can be proved that if we set
\begin{align*}
  X=&\Bigl\{(S_0, E_0, I_0,V_0, S_1, E_1, I_1,V_1, S_h, E_h, I_h,V_h )\in\mathbb{R}_+^{12}~\vert\\
  &E_0, I_0, E_1, I_1, E_h, I_h,V_h\in\mathbb{R}_+\cup\{0\}\Bigr\}\\
\end{align*}
the region
\begin{align*}
\Gamma =&\Bigl\{p\in X~\vert~S_0 + E_0 + I_0 + V_0 + S_1 + E_1 + I_1 +\\
& V_1 + S_h + E_h +I_h+V_h\le\frac{A+H}{\min\{\mu + c, d,\mu_h\}}\Bigr\}\\
\end{align*}
is positively invariant.

It is proved that there are two equilibria 
\begin{itemize}
	\item a disease free equilibrium
	$E_0=(S_{0,0},0,0,V_{0,0},S_{1,0},0,0,$ $V_{1,0},S_{h,0},0,0,0)$ where the
	complete expression is reported in \cite{paperNostro};
	\item an endemic equilibrium $E*$ whose expression is  to complicate to
	formally compute, which can be proved to exist if $R_0>1$\cite{papernostro}.
\end{itemize}

After having computed $R_0$, whose expression is presented in
\cite{paperNostro}, it can be proved that if $R_0<1$, $E_0$ is globally
asymptotically stable for system \ref{s4}.

Numerical simulations can be performed using parameters estimated from data
provided by the Department of Health of Guangdong Province. Fig. \ref{fig:dsh1}
and Fig. \ref{fig:dsh2} show a simulation of human rabies infection cases over
time for the Guangdong Province.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}Simulation of human rabies infection cases from 2006 to 2014 for the Guangdong Province of China. This plot reproduces Fig. 3.2 of \\cite{Cina}."}
source("SEIR_dogs_stray_humans.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme +
  ylab(TeX("$I_h$")) + xlab("time (months)") + xlim(0, 100)
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh2}Numerical simulations of rabies cases in the Guangdong Province from 2006 to 2014. $S(t)$ and $V(t)$ are plotted in log$_{10}$ scale."}
ggplot(data = SEIR_out_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab(latex2exp::TeX("$E_h$, $I_h$")) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~., name =
      latex2exp::TeX("$S_h$, $V_h$"),
    breaks = breaks, labels = labels)) + theme(legend.position="bottom",
      axis.title.y.right = element_text(margin = margin(l = 20))) + 
  xlab("time (months)")
```

We can examine how different values of the parameters influence the dynamic of
$I(t)$. Fig. \ref{fig:dsh3} shows the dynamic of $I_h(t)$ for different values
of $l$, the rate of domestic dog abandonment, and $\epsilon$, the rate of
infectious domestic dog abandonment. We can observe that reducing $\epsilon$ is
more effective than reducing $l$.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh3}Simulations of infected human rabies cases $I_h(t)$ in the Guangdong Province for different values of $l$ and $\\epsilon$. This plot reproduces Fig. 3.5a of \\cite{Cina}."}
ggplot(data = I_h_df_long_1, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$l$, $\\epsilon$"), labels = labels_1) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)") + xlim(0, 125)
```

Fig. \ref{fig:dsh4} shows the dynamic of $I_h(t)$ for different values of $c$,
the culling rate of stray dogs, and $\nu$, the vaccination rate of domestic
dogs. We can observe that an increasing the vaccination rate is more effective
in reducing $I_h(t)$ than increasing the culling rate.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh4}Simulations of infected human rabies cases $I_h(t)$ in the Guangdong Province for different values of $c$ and $\\nu$. This plot reproduces Fig. 3.5b of \\cite{Cina}."}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "", labels = c(
    "No intervention", "50% stray dog culling", "100% stray dog culling",
    "50% vaccination", "70% vaccination")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + guides(col = guide_legend(nrow = 3)) + xlim(0, 125)
```

Finally, Fig. \ref{fig:dsh5} shows the dynamic of $I_h(t)$ for different values
of $\nu_h$, the culling rate of stray dogs, the vaccination rate of exposed
humans. Increasing $\nu_h$ decreases $I_h(t)$, but does not stop the epidemic.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh5}Simulations of infected human rabies cases $I_h(t)$ in the Guangdong Province for different values of $\\nu_h$. This plot should reproduces Fig. 3.5c of \\cite{Cina}, but is quite different."}
ggplot(data = I_h_df_long_3, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$\\nu_h$"), labels = nu_h_vect) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + xlim(0, 125)
```

## Seasonal SEIR model

It is well-known that many diseases exhibit seasonal fluctuations. A number of
phenomenons can be regarded as sources of periodicity, for instance seasonal
changes in transmission rate and birth rate. The previous models failed to grasp
this periodic component of the rabies epidemic. In this section, a new SEIR
model that highlights seasonal changes in the transmission rate is examined. The
model can be written as:
\begingroup
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s5}
	\begin{cases}
		S_d'(t) = A + \lambda_dR_d + \sigma_d (1 - \gamma_d )E_d - \beta_d(t) S_dI_d - (m_d + k_d)S_d\\
		E_d'(t) = \beta_d(t) S_dI_d - \sigma_d (1 - \gamma_d )E_d - \sigma_d\gamma_d E_d - (m_d + k_d)E_d\\
		I_d'(t) = \sigma_d\gamma_d E_d - (m_d + \mu_d)I_d\\
		R_d'(t) = k(S_d + E_d) - (m_d + \lambda_d)R_d\\
		S_h'(t) = B + \lambda_hR_h + \sigma_h(1 - \gamma_h )E_h - m_hS_h - \beta_{dh}(t)S_hI_h\\
		E_h'(t) = \beta_{dh}(t)S_hI_d - \sigma_h(1 - \gamma_h )E_h - \sigma_h\gamma_hE_h - (m_h + k_h )E_h\\
		I_h'(t) = \sigma_h\gamma_hE_h - (m_h + \mu_h)I_h\\\
		R_h'(t) = k_hE_h - (m_h + \lambda_h )R_h
	\end{cases}
\end{align}
\endgroup

where the transmission rates are $\beta_d(t)=a[1 + bsin( \frac{\pi}{6} t + 5.5)]$
and $\beta_{dh}(t)=a_1[1 + b_1sin( \frac{\pi}{6} t + 5.5)]$. Even in this
case, if we set
$$
X=\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in (\mathbb{R}_+\cup\{0\})^8\Bigr\}
$$

it can be proved that there is a positively invariant region
\begin{align*}
\Gamma =&\Bigl\{p\in X~\vert~0 < S_d + E_d + I_d + R_d \le \frac{A}{m_d}\text{ and }\\
&0 < S_h + E_h + I_h + R_h \le \frac{B}{m_h}\Bigr\}\\
\end{align*}
It is proved that the system \ref{s5} has 
\begin{itemize}
	\item a disease free equilibrium $E_0=(S_{0},0,0,R_{0},S_{h},0,0,0)$ where the
	complete expression is reported in \cite{paperNostro}, which is globally
	asymptotically stable for $R_0<1$;
	\item at least one positive periodic solution \cite{111paperNostroBio}.
\end{itemize}

Fig. \ref{fig:s1} shows a numerical simulation of the model for the number of
human rabies cases in China starting in 2004. Data from the Chinese Ministry of
Health was used to estimate the parameters. Fig. \ref{fig:s2} shows the same
simulation for the 4 SEIV classes. From these simulation we can see a reduction
in the number of human rabies cases in the short future. However the epidemic is
not eradicated, as we can see from a simulation of the same model for a period
of 125 years (Fig. \ref{fig:s3}).

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s1}Simulation of human rabies infection cases in China from 2004, for a period of 12.5 years. This plot reproduces Fig. 4.2a of \\cite{Cina}."}
source("SEIR_seasonal.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
  ylab(TeX("$I_h$")) + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s2}Numerical simulations of human rabies cases in China from 2004, for a period of 12.5 years. $S(t)$ and $R(t)$ are plotted in log$_{10}$ scale."}
ggplot(data = SEIR_out_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab(latex2exp::TeX("$E_h$, $I_h$")) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~., name =
      latex2exp::TeX("$S_h$, $R_h$"),
    breaks = breaks, labels = labels)) + theme(legend.position="bottom",
      axis.title.y.right = element_text(margin = margin(l = 20))) + 
  xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s3}Simulation of human rabies infection cases in China from 2004, for a period of 125 years. This plot reproduces Fig. 4.2b of \\cite{Cina}."}
ggplot(data = as.data.frame(SEIR_out_2)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
  ylab(TeX("$I_h$")) + xlab("time (months)")
```

We can examine how the initial conditions influence the dynamic of $I(t)$. The
initial number of susceptible dogs has a strong influence on the number of human
rabies cases, as shown in Fig. \ref{fig:s4}. This underlines the importance of
controlling the dog population in order to mitigate the epidemic. Contrarily,
the initial number of infectious dogs has a very small effect on the number of
human rabies cases, as shown in in Fig. \ref{fig:s5}.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s4}Number of human rabies cases $I_h(t)$ for different initial values of the susceptible dog population $S_d(0)$. This plot reproduces Fig. 4.3a of \\cite{Cina}."}
ggplot(data = I_h_df_long_1, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$S_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + guides(col = guide_legend(nrow = 2))
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s5}Number of human rabies cases $I_h(t)$ for different initial values of the infectious dog population $I_d(0)$. This plot reproduces Fig. 4.3b of \\cite{Cina}."}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$I_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)")
```

## Multi-patch SEIV model

In 2005 the Chinese government decided to monitor rabies at the national level
in an attempt to obtain a more comprehensive epidemiological dataset. The data
highlight that east China is epidemiologically related to adjoining provinces
and to distant provinces. Moreover it has been demonstrated that there is a
general trend of longitudinal transmission and latitudinal transmission
\cite{106, 108}.
	
A new SIEV model is developed to analyze how the movement of dogs affects the
spread of rabies. Reaction-diffusion equations are based on the assumptions that
the spatial domain is connected and the movement of dogs is a continuous process
in the domain. These assumptions cannot be considered satisfied in China. Thus,
in order to include the spatial aspect, it is preferred to divide China in
smaller areas, called patches, generating a multi-patch model. This is an ODE
model in which two travel rate matrices $\Phi$ and $\Psi$ are used to model the
movement of each class of dogs and humans from a patch to another. The resulting
model is:
\begingroup
\small
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s6}
	\begin{cases}
	S_{d,i}'(t) = & A_i + \lambda_{d,i}V_{d,i} + \sigma_{d,i} (1 - \gamma_{d,i})E_{d,i} - \beta_{d,i}S_{d,i}I_{d,i} +\\
	& - (m_{d,i} + k_{d,i})S_{d,i}+\sum_{j=1}^n\Phi^S_{ij}S_{d,j}\\
	E_{d,i}'(t) = & \beta_{d,i}  S_{d,i}I_{d,i} - \sigma_{d,i} (1 - \gamma_{d,i}
	)E_{d,i} - \sigma_{d,i}\gamma_{d,i} E_{d,i} +\\
	& - (m_{d,i} + k_{d,i})E_{d,i}+\sum_{j=1}^n\Phi^E_{ij}E_{d,j}\\
	I_{d,i}'(t) = & \sigma_{d,i}\gamma_{d,i} E_{d,i} - (m_{d,i} +
	\mu_{d,i})I_{d,i}+\sum_{j=1}^n\Phi^I_{ij}I_{d,j}\\
	V_{d,i}'(t) =& k_{d,i}(S_{d,i} + E_{d,i}) - (m_{d,i} + \lambda_{d,i})V_{d,i}
	+\sum_{j=1}^n\Phi^V_{ij}V_{d,j}\\
	S_{h,i}'(t) =& B + \lambda_{h,i}V_{h,i} + \sigma_{h,i}(1 - \gamma_{h,i} )E_{h,i}
	- m_{h,i}S_{h,i} +\\
	& - \beta_{dh,i}S_{h,i}I_{d,i}+\sum_{j=1}^n\Psi^S_{ij}S_{h,j}\\
	E_{h,i}'(t) =& \beta_{dh,i} S_{h,i}I_{d,i} - \sigma_{h,i}(1 - \gamma_{h,i}
	)E_{h,i} - \sigma_{h,i}\gamma_{h,i}E_{h,i} +\\
	& - (m_{h,i} + k_{h,i} )E_{h,i}	+\sum_{j=1}^n\Psi^E_{ij}E_{h,j}\\
	I_{h,i}'(t) =& \sigma_{h,i}\gamma_{h,i}E_{h,i} - (m_{h,i} + \mu_{h,i})I_{h,i} +
	\sum_{j=1}^n\Psi^I_{ij}I_{h,j}\\
	V_{h,i}'(t) =& k_{h,i}E_{h,i} - (m_{h,i} + \lambda_{h,i}
	)V_{h,i}+\sum_{j=1}^n\Psi^V_{ij}V_{h,j}\\
	\end{cases}
\end{align}
\endgroup

This model consists of a classical SEIV model for every patch $i$ for
$i\in\{1,\dots,n\}$, plus the migration component. It is important to underline
that 
$$
\Phi^X_{ii}=-\sum_{\substack{j=1 \\ j\neq i}}^n\Phi^X_{ij} \quad\text{and}\quad\Psi^X_{ii}=-\sum_{\substack{j=1 \\ j\neq i}}^n\Psi^X_{ij} 
$$

for $i\in\{1,\dots,n\}$ and for $X\in\{S,E,I,V\}$. Moreover, $\Phi^X_{ij}\ge0$
and $\Psi^X_{ij}\ge0$ for $i,j\in\{1,\dots,n\}$ and $X\in\{S,E,I,V\}$.

It can be proved that there exist a disease free equilibrium
$E_0=(\underline{S}_{d,0},\underline{0},\underline{0},\underline{V}_{d,0},\underline{S}_{h,0},\underline{0},\underline{0},\underline{0})$,
whose expression is presented in \cite{Cina}.

Even trying to compute the formal expression of $R_0$ is extremely complex for
two patches. Therefore the the study of this model behavior is based on the
numerical simulations.

In \cite{Cina} the initial conditions of numerical simulations for multi-patch
models are not reported. Thus, we cannot fully reproduce the original results.
However, we can still use plausible values of the initial conditions to observe
the qualitative behavior of the system. We focus on a two-patches model for the
provinces of Hebei and Fujian. It is estimated that $R^{Hebei}_0=0.5477$ and
$R^{Fujian}_0=0.8197$, so the disease would die out in both provinces since
their isolated basic reproduction numbers are less than one. However, the
immigration of dogs from Fujian to Hebei leads to the persistence of rabies in
this province, as can be seen in Fig. \ref{fig:p1}. This example theoretically
shows the possibility that the immigration of dogs can lead the disease to a
worse scenario even it could be eliminated in each isolated patch.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:p1}Simulations of the numbers of reported human rabies cases in Hebei and Fujian from 2004 to 2012. This plot attempts to reproduces Fig. 6.3 of \\cite{Cina}, using plausible values of the initial conditions."}
source("SEIV_patch.R")

ggplot(data = out, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "Province", labels =
      c("Hebei", "Fujian")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + xlim(0, 100)
```

# Novel results

Here we present novel results obtained using two new models: a multi-patch SEIV
model including stray dogs and a multi-patch SEIV model including stray dogs and
seasonal effects. In these models we study the effects of parameters on the
diffusion of rabies. In particular we foucs on three parameters: the culling
rate of stray dogs, the vaccination rate of dogs and the fertility of stray
dogs.

## Multi-patch SEIV model with stray dogs

First, we present a SEIV model that takes into account the distinction between
stray and domestic dogs. This model is similar to the one discussed previously,
with the addition of terms accounting for the abandonment rate of domestic dogs
and  the transmission of rabies between stray dogs and domestic dogs or humans.
The flow chart of the model is reported in Fig. \ref{fig:mess} and the system of
ODEs for the model is:

\begingroup
\small
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s6}
	\begin{cases}
  S_{s,i}'(t) = & A_{s,i} + \lambda_{s,i}V_{s,i} + \sigma_{s,i} (1 - \gamma_{s,i})E_{s,i} - \beta_{ss,i}S_{s,i}I_{s,i} +\\
	& - \beta_{ds,i}S_{s,i}I_{d,i} - (m_{s,i} + k_{s,i})S_{s,i}+\sum_{j=1}^nP^S_{ij}S_{s,j} + lS_{d,i}\\
	E_{s,i}'(t) = & \beta_{ds,i}  S_{s,i}I_{d,i} + \beta_{ss,i}  S_{s,i}I_{s,i} + lE_{d,i} - \sigma_{s,i} (1 - \gamma_{s,i}
	)E_{s,i} \\
	& - \sigma_{s,i}\gamma_{s,i} E_{s,i} - (m_{s,i} + k_{s,i})E_{s,i}+\sum_{j=1}^nP^E_{ij}E_{s,j} + \\
	I_{s,i}'(t) = & \sigma_{s,i}\gamma_{s,i} E_{s,i} - (m_{s,i} +
	\mu_{s,i})I_{s,i}+\sum_{j=1}^nP^I_{ij}I_{s,j} + \epsilon I_{d,i}\\
	V_{s,i}'(t) =& k_{s,i}(S_{s,i} + E_{s,i}) - (m_{s,i} + \lambda_{s,i})V_{s,i} +\sum_{j=1}^nP^V_{ij}V_{s,j} + lI_{d,i}\\
	S_{d,i}'(t) = & A_{d,i} + \lambda_{d,i}V_{d,i} + \sigma_{d,i} (1 - \gamma_{d,i})E_{d,i} - \beta_{dd,i}S_{d,i}I_{d,i} +\\
	& -\beta_{ds,i}S_{d,i}I_{s,i} - (m_{d,i} + k_{d,i})S_{d,i}+\sum_{j=1}^n\Phi^S_{ij}S_{d,j} -  lS_{d,i}\\
	E_{d,i}'(t) = & \beta_{dd,i}  S_{d,i}I_{d,i} + \beta_{ds,i}  S_{d,i}I_{s,i} - lE_{d,i} - \sigma_{d,i} (1 - \gamma_{d,i}
	)E_{d,i} +\\
	& - \sigma_{d,i}\gamma_{d,i} E_{d,i} - (m_{d,i} + k_{d,i})E_{d,i}+\sum_{j=1}^n\Phi^E_{ij}E_{d,j}\\
	I_{d,i}'(t) = & \sigma_{d,i}\gamma_{d,i} E_{d,i} - (m_{d,i} +
	\mu_{d,i})I_{d,i}+\sum_{j=1}^n\Phi^I_{ij}I_{d,j} - \epsilon I_{d,i}\\
	V_{d,i}'(t) =& k_{d,i}(S_{d,i} + E_{d,i}) - (m_{d,i} + \lambda_{d,i})V_{d,i} +\sum_{j=1}^n\Phi^V_{ij}V_{d,j} -  lI_{d,i}\\
	S_{h,i}'(t) =& B + \lambda_{h,i}V_{h,i} + \sigma_{h,i}(1 - \gamma_{h,i} )E_{h,i}
	- m_{h,i}S_{h,i} +\\
	& - \beta_{hd,i}S_{h,i}I_{d,i} - \beta_{hs,i}S_{h,i}I_{s,i} + \sum_{j=1}^n\Psi^S_{ij}S_{h,j}\\
	E_{h,i}'(t) =& \beta_{dh,i} S_{h,i}I_{d,i} - \sigma_{h,i}(1 - \gamma_{h,i}
	)E_{h,i} - \sigma_{h,i}\gamma_{h,i}E_{h,i} +\\
	& - (m_{h,i} + k_{h,i} )E_{h,i}	+\sum_{j=1}^n\Psi^E_{ij}E_{h,j}\\
	I_{h,i}'(t) =& \sigma_{h,i}\gamma_{h,i}E_{h,i} - (m_{h,i} + \mu_{h,i})I_{h,i} +
	\sum_{j=1}^n\Psi^I_{ij}I_{h,j}\\
	V_{h,i}'(t) =& k_{h,i}E_{h,i} - (m_{h,i} + \lambda_{h,i}
	)V_{h,i}+\sum_{j=1}^n\Psi^V_{ij}V_{h,j}\\
	\end{cases}
\end{align}
\endgroup

```{r, include=FALSE}
wdgt <- DiagrammeR::grViz("patch_SEIV.gv")
wdgt_html_tf <- tempfile(fileext = ".html")
htmlwidgets::saveWidget(wdgt, wdgt_html_tf)
webshot::webshot(url = sprintf("file://%s", wdgt_html_tf), 
  selector = "#htmlwidget_container",
  file = "patch_SEIV.pdf")
unlink(wdgt_html_tf)
```
\begin{figure*}[!htb]
  \centering
  \includegraphics[width=\textwidth,trim = 4cm 7.1cm 2.5cm 1.2cm, clip]{patch_SEIV.pdf}
  \caption{Flow chart of the SEIV model with stray and domestic dogs. The meaning of parameters is identical to the multi-patch model presentes in \cite{Cina}. Additional parameters not present in that model are: the transmission rated between the different classes $\beta_{dd}$, $\beta_{ds}$, $\beta_{ss}$, $\beta_{hd}$, $\beta_{hs}$,}; the abandonment rates of infectious and non-infectious domestic dog $\epsilon$ and $l$, respectively; the travel rate matrices of stray dogs $P^J$, $J=S,E,I,V$. 
  \label{fig:mess}
\end{figure*}

For this model we use parameters taken form the previous models, in particular
the multi-patch SEIV model for the provinces of Hebei and Fujian and the SEIV
model with stray and domestic dogs. Furthermore, we assume that the Fujian
province has a larger number of stray dogs than the Hebei province at $t=0$.

Fig. \ref{fig:ps1} shows...

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:ps1}plotting example"}
source("SEIV_patch_stray.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h_1), color = "blue") + 
  geom_line(mapping = aes(time, I_h_2), color = "red") + myTheme
```

## Multi-patch seasonal SEIV model

Here we present an extension of the previous model, that takes into account the
seasonal effect considered in \cite{Cina}. The structure of the model is
identical to the one discussed in the previous section. In this model, the
transmission rates are functions of time:
$$\beta_{dd}(t)=\beta_{ss}(t)=\beta_{ds}(t)=a[1 + bsin( \frac{\pi}{6} t + 5.5)]$$
$$\beta_{hd}(t)=\beta_{hs}(t)=a_1[1 + b_1sin( \frac{\pi}{6} t + 5.5)]$$
For simplicity we assume that the transmission rates are unaffected by the dogs
being stray or domestic. Taking this difference into account may result in a
more realistic model, but would require additional parameters that we cannot
estimate with the data available.

Fig. \ref{fig:pss1} shows...

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:pss1}plotting example"}
source("SEIV_patch_stray_seasonal.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h_1), color = "blue") + 
  geom_line(mapping = aes(time, I_h_2), color = "red") + myTheme
```

## Parameters

The current literature suggests that culling of stray dogs in order to reduce
the dg population is an uneffective and insostenible strategy. New approaches to
reduce the dog population have been proposed, such as the use of
immunocontraceptive drugs, like GonaCon, a synthetic gonadotrophin-releasing
hormone (GnRH) which is able to induce long-term reduction of fertility in both
genders. It has also been proposed that GnRH could be administered together with
rabies vaccine in a dual immunocontraceptive vaccine, which would serve as an
ethical and highly efficacious mean of both controlling dog populations and
protecting against rabies. The principal advantage of a dual immunocontraceptive
vaccine is that it might be suitable for oral administration, allowing its
administration via bait \cite{Franka}.

In this section, we examine the effect of three parameters on the diffusion of
rabies: the culling rate of stray dogs, the vaccination rate of dogs and the
fertility of stray dogs. In particular, we want to use our models to analyze
what would be the effect of the administration of a dual immunocontraceptive
vaccine.

# Conclusions

Different models have been proposed and summarized to describe the dynamic of
the rabies epidemic in China. However some small inconsistencies in \cite{Cina}
has made studying this paper challenging. Moreover data deficit has made
impossible to estimate the accuracy of the new proposed patches model.

\begin{thebibliography}{1}
	
	\bibitem{Wang}Wang L, Tang Q, Liang G. Rabies and rabies virus in wildlife in
	mainland China, 1990-2013. Int J Infect Dis. 2014; 25:122-9.
	
	\bibitem{Schnell}Schnell MJ, McGettigan JP, Wirblich C, Papaneri A. The cell
	biology of rabies virus: using stealth to reach the brain. Nat Rev Microbiol.
	2010; 8(1):51-61.
	
	\bibitem{Velasco}Velasco-Villa A, Escobar LE, Sanchez A, Shi M, Streicker DG,
	Gallardo-Romero NF, Vargas-Pino F, Gutierrez-Cedillo V, Damon I, Emerson G.
	Successful strategies implemented towards the elimination of canine rabies in
	the Western Hemisphere. Antiviral Res. 2017; 143:1-12.
	
	\bibitem{Hemachudha}Hemachudha T, Ugolini G, Wacharapluesadee S, Sungkarat W,
	Shuangshoti S, Laothamatas J. Human rabies: neuropathogenesis, diagnosis, and
	management. Lancet Neurol. 2013; 12(5):498-513.
	
	\bibitem{Cina} 
	Ruan S.Modeling the transmission dynamics and control of rabies in China.
	Mathematical biosciences 286. 2017; 65-93.
	
	\bibitem{54} 
	J. Ma , Z. Ma. Epidemic threshold conditions for seasonally forced SEIR models.
	Math. Biosci. Eng. 3. 2006; 161 -172.
	 
	\bibitem{78} 
	I. Schwartz. Small amplitude, long periodic out breaks in seasonally driven
	epidemics. J. Math. Biol. 30. 1992; 473 -491 .
	
	\bibitem{113} 
	J. Zhang, Z. Jin, G. Sun, T. Zhou, S. Ruan. Analysis of rabies in China:
	transmission dynamics and control. PLoS One 6. 2011; e20891.
	
	\bibitem{111} 
	J. Zhang, Z. Jin, G. Sun, T. Zhou, S. Ruan. Analysis of rabies in China:
	transmission dynamics and control. PLoS One 6. 2011; e20891.
	
	\bibitem{108} 
	J. Yu, L. Hao, Q. Tang. The spatial and temporal dynamics of rabies in China.
	PLoS Negl. Trop. Dis. 6. 2012; e1640.
	
	\bibitem{106} 
	C. Yin, H. Zhou, H. Wu. Analysis on factors related to rabies epidemic in China
	from 2007 -2011. Virol. Sin. 27. 2012; 132 -143.
	
	\bibitem{23} 
	Diekmann, J.A.P. Heesterbeek, M.G. Roberts. The construction of next-generation
	matrices for compartmental epidemic models. J. R. Soc. Interface 7. 2010; 873
	-885.
	
	\bibitem{96} 
	X. Wang, J. Lou. Two dynamic models about rabies between dogs and human. Biol.
	Syst. 16. 2008; 519 -529.
	
	\bibitem{mate1} 
	Stefan Ma, Yingcun Xia. Mathematical Understanding of Infectious Disease
	Dynamics. Lecture Notes Series, Institute for Mathematical Sciences, National
	University of Singapore: Volume 16. 2008; 33-35.
	
	\bibitem{Franka} Franka R, Smith TG, Dyer JL, Wu X, Niezgoda M, Rupprecht CE.
	Current and future tools for global canine rabies elimination. Antiviral Res.
	2013; 100(1):220-5.

\end{thebibliography}