---
title: SIER models for rabies epidemic in China
author: Martina Iannacito\textsuperscript{1} and Matteo Ciciani\textsuperscript{2}
affiliation:
  - \textsuperscript{1}\textit{...}
  - \textsuperscript{2}\textit{Master in Quantitative and Computational Biology, University of Trento, Italy}
output:
    pdf_document:
      latex_engine: xelatex
      template: SelfArx-template.tex
abstract: Here is the abstract \cite{Cina}
---
# Introduction

## The biology of rabies
Rabies is one of the oldest diseases in recorded human history, dating back more
than 4000 years. More than 70000 human rabies deaths occur annually, and Asia
accounts for approximately 80\% of the worldwide total \cite{Wang}. As 40\% of
those infected are children, the years of life lost by this infection makes
rabies the seventh most important global infectious disease \cite{Schnell}.
Currently, China is second only to India in the number of people killed by
rabies worldwide, and the epidemic area has expanded to almost the entire
country \cite{Wang}.

Rabies is an acute encephalomyelitis (inflammation of the brain and the spinal
cord) caused by viruses in the genus \emph{Lyssavirus} of the family
\emph{Rhabdoviridae}. Rabies lyssavirus (RABLV), the prototype species of the
genus, causes disease in nearly all terrestrial mammals. Despite the worldwide
distribution of RABLV, only a few animal species in the orders \emph{Carnivora},
\emph{Chiroptera} (bats) and \emph{Primates} play a critical role in maintaining
the circulation of the virus in nature. Most of these mammals harbor
species-specific RABLV variants which circulate within discrete geographic
boundaries. However, because almost all cases of human rabies result from the
bites of rabid dogs, dog-maintained viruses are by far the most important target
of control measures \cite{Velasco}.

As for most viruses, the rabies virus life cycle can be divided into three
phases. The first phase includes entry into the host cell by endocytosis
followed by release of the viral genome into the cytoplasm. The classical
wild-type rabies virus first infects a muscle cell, then enters a motor neuron
at the neuromuscular junction. Rabies virus particles are subsequently
transported in a retrograde direction through the axon of the infected neuron.
Once the virus particles reach the cell body of the neuron, the second phase
commences, which involves the production of the virion components. The last
phase of the life cycle comprises the assembly of the viral components and the
release of mature rabies virus particles, which can then start a new round of
infection \cite{Schnell}.

By the time a patient develops the first prodromal symptoms, such as fever,
flu-like symptoms, and gastrointestinal disturbances, the virus is already
widely disseminated throughout the central nervous system \cite{Hemachudha}.

Clinical manifestation of rabies in humans has two forms: the furious form (80\%
of infections), and the paralytic form (20\% of infections). The furious form of
rabies is characterized by hydrophobia, episodes of hallucinations and
excitement. These behavioral changes occur simultaneously with the shedding of
large amounts of rabies virus in the saliva, which facilitates the spread of the
virus to a new host. The paralytic form of rabies is characterized by weakness
and flaccid paralysis. In both cases, survival after the onset of symptoms is
rarely more than 7 days and the disease is lethal in almost 100\% of cases
\cite{Schnell}.

Human rabies virus infections are controlled by limiting the infection in
animals and by post-exposure prophylaxis (PEP) of humans after contact with a
potentially infected animal. Current PEP for previously non-immunized
individuals consists of one dose with rabies immunoglobulin in conjunction with
five doses of immunization with inactivated rabies virus vaccines. This regimen
is effective, but its usefulness in developing countries, where rabies is
endemic, is hampered by the high cost and lack of compliance \cite{Schnell}.

## SEIR models

One of the most known model used to investigate the evolution of epidemics is
SEIR. It is coherent to apply a basic SEIR model if the disease considered has
an average duration shorter than the average time needed to observe significant
modifications in the population structure. For example it can be applied to
describe flu epidemics in the human species, since the average lifespan of a
human being and the time needed to generate a new individual are on average
longer than the duration of the flu. Under this assumption demographical aspects
are negligible and consequently the model focuses only on the disease dynamic.
In order to apply the model the total population $N(t)$ at time t is split into
four classes:
\begin{itemize}
  \item $S(t)$ : the class of individuals susceptible to the disease at time t;
  \item $E(t)$ : the class of individuals exposed to the pathogen and not
  contagious at time t;
  \item $I(t)$ : the class of infectious individuals at time t;
  \item $R(t)$ : the class of removed individuals, i.e. those who get out of the
  model because they are dead or they have become immune at time t.
\end{itemize}

Then it is assumed that susceptible individuals who come into contact with
infected ones become exposed at rate $\lambda(t)$ , exposed move into the
infectious class at rate $k(t)$ and at last infectous are removed at rate
$\gamma(t)$ . The force of infection $\lambda(t)$ is mathematically described
as:

$$
\lambda(t) = c(t)\chi\frac{I(t)}{N(t)}
$$

where $c(t)$ is the average number of contact of one individual in the unit of
time, $\chi$ is the probability to become infect after a contact with an
infectious and $\frac{I(t)}{N(t)}$ is the total number of infectious. To make
computations easier it is assumed that all these parameter are constant in time
and thanks to the absence of demographic aspects the total population is also
constant. Therefore the SEIR ODE system is:

\begin{align}
  \label{s1}
  \begin{cases}
    S'(t) = -\frac{c\chi}{N}S(t)I(t)\\
    E'(t) = \frac{c\chi}{N}S(t)I(t)-kE(t)\\
    I'(t) = kE(t)-\gamma I(t)\\
    R'(t) =\gamma I(t)
  \end{cases}
\end{align}

It is important to underline that if $\gamma$ is the rate at which an infectious
is removed, then $\gamma$ is the average time an individual spends as
infectious. Consequently the basic epidemiologic reproduction number is the
number of ill individuals generated by an infectious in the life time as
infectious, i.e.

$$
R_0 = \frac{c\chi}{\gamma}
$$

It is possible to obtain a simpler system rescaling \ref{s1} in the following
way: $\tau = \gamma t$, $u(t) =\frac{S(t)}{N}$, $v(t) = \frac{E(t)}{N}$, $w(t) =
\frac{I(t)}{N}$ and $z(t) =\frac{R(t)}{N}$. Thanks to the applied rescaling,
it follows that $u+v+w+z=\frac{S+E+I+R}{N}=1$. Denoting
$\dot{u}=\frac{du}{d\tau}$, the system (\ref{s1}) becomes:
\begin{align*}
	\begin{cases}
		\dot{u}(\tau) =-\frac{c\chi}{\gamma}u(\tau)w(\tau)\\
		\dot{v}(\tau) = \frac{c\chi}{\gamma}u(\tau)w(\tau)-\rho v(\tau)\\
		\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
		\dot{z}(\tau) = w(\tau)
	\end{cases}
\end{align*}

which is equivalent to
\begin{align}
	\label{s2}
	\begin{cases}
		\dot{u}(\tau) = -R_0u(\tau)u(\tau)w(\tau)\\
		\dot{v}(\tau) = R_0u(\tau)w(\tau)-\rho v(\tau)\\
		\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
		\dot{z}(\tau) = w(\tau).
	\end{cases}
\end{align}

It can be proved as in \cite{mate1} that given positive initial condition for
the system (\ref{s2}) there is a unique positive solution.

In order to make the model more realistic, it is convenient to introduce a
simple vital dynamic with death rate $\mu$ equal to birth rate $\beta$. Under
this hypothesis the population is still constant in time. The resulting system
is:
\begin{align}
  \label{s3}
	\begin{cases}
    \dot{u}(\tau) = -R_0u(\tau)u(\tau)w(\tau)+\beta(1-u(\tau))\\
    \dot{v}(\tau) = R_0u(\tau)w(\tau)-(\rho+\beta)v(\tau)\\
    \dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
    \dot{z}(\tau) = w(\tau)-\beta z.
  \end{cases}
\end{align}

Reminding that $u+v+z+w=1$, it is easy to compute the equilibria of the system
(\ref{s3}), which are:
\begin{itemize}
	\item the disease free equilibrium $E_0=(1,0,0,0)$;
	\item the endemic equilibrium $E^*=(u^*,v^*,w^*,z^*)$, that exists if and only
	if holds the condition
	\begin{equation}
  	\label{c1}
  	R_0>1+\frac{\beta}{\rho}
	\end{equation}
\end{itemize}

To investigate the stability of the equilibria, it is useful to consider the
jacobian of the reduced system:
\begin{align*}
	\begin{cases}
  	\dot{u}(\tau)=-R_0u(\tau)u(\tau)w(\tau)+\beta(1-u(\tau))\\
  	\dot{v}(\tau) = R_0u(\tau)w(\tau)-(\rho+\beta)v(\tau)\\
  	\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
	\end{cases}
\end{align*}

For the disease free equilibrium the Jacobian is
$$
J(E_0)=\begin{bmatrix} 
-\beta & 0  & -R_0 \\ 
0 &	-(\rho+\beta)	& R_0\\
0 &	\rho	& -1\\
\end{bmatrix}
$$

It can be observed that it is a block matrix. Consequently an eigenvalue is
$\lambda_1=-\beta$, the other ones are the eigenvalues of the matrix
$$
J(E_0)=\begin{bmatrix} 
-(\rho+\beta)	& R_0\\
\rho	& -1\\
\end{bmatrix}
$$
which has negative trace and positive determinant if the condition (\ref{c1})
is false. For the endemic equilibrium the Jacobian is
$$
J(E^*)=\begin{bmatrix} 
-(R_0w^*+\beta) & 0  & -R_0u^* \\ 
R_0w^* &	-(\rho+\beta)	& R_0u^*\\
0 &	\rho	& -1\\
\end{bmatrix}
$$
Using the Routh-Hurwitz conditions, it can be proved as in \cite{mate1} that the
endemic equilibrium, when it exists, is asymptotically stable.

Numerical simulation for the a genericl SEIR model are shown in Fig.
\ref{fig:gen1} and \ref{fig:gen2}, for values of $R_0>1$ and $R_0<1$,
respectively.

```{r, echo=FALSE, fig.width=7, fig.height=6,fig.cap="\\label{fig:gen1}Numerical simulation of a SEIR model with $R_0>1$. We can see that the system convergest to the endemic equilibrium."}
source("SEIR_general.R")

ggplot(data = SEIR_out_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab("E, I, R") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*2, name = "S")) + 
  theme(legend.position="bottom", axis.title.y.right =
      element_text(margin = margin(l = 20))) + 
  xlab("time")
```

```{r, echo=FALSE, fig.width=7, fig.height=6,fig.cap="\\label{fig:gen2}Numerical simulation of a SEIR model with $R_0<1$. We can see that the system convergest to the disease-free equilibrium."}
source("SEIR_general.R")

ggplot(data = SEIR_out_long2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab("E, I, R") +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*10, name = "S")) + 
  theme(legend.position="bottom", axis.title.y.right =
      element_text(margin = margin(l = 20))) + 
  xlab("time")
```

The basic structure of the SEIR system can be adequately modified and used to
describe the rabies epidemic behavior as in \cite{PAPERNOSTRO?}.

# Reproduction of original results

## SEIR model for dog-human interactions

As it has been pointed out in the biological introduction to rabies, following a
bite from an infectious individual, the virus has an incubation period in the
new host. Consequently the bitten host can be considered exposed. It is
therefore appropriate to describe this epidemic phenomena through a SEIR model
for the dogs and the human population. The most important difference with the
classical SEIR is that in this model only the exposed individuals that are
subjected to PEP can move into the removed class. Moreover, the removed can
loose immunity and get back to susceptible class. The ODE system is then:

\begingroup
\setlength{\mathindent}{0pt}
\begin{align*}
	\begin{cases}
		S_d'(t) = A + \lambda_dR_d + \sigma_d (1 - \gamma_d )E_d - \beta_d S_dI_d - (m_d + k_d)S_d\\
		E_d'(t) = \beta_d S_dI_d - \sigma_d (1 - \gamma_d )E_d - \sigma_d\gamma_d E_d - (m_d + k_d)E_d\\
		I_d'(t) = \sigma_d\gamma_d E_d - (m_d + \mu_d)I_d\\
		R_d'(t) = k(S_d + E_d) - (m_d + \lambda_d)R_d\\
		S_h'(t) = B + \lambda_hR_h + \sigma_h(1 - \gamma_h )E_h - m_hS_h - \beta_{dh}S_hI_h\\
		E_h'(t) = \beta_{dh}S_hI_d - \sigma_h(1 - \gamma_h )E_h - \sigma_h\gamma_hE_h - (m_h + k_h )E_h\\
		I_h'(t) = \sigma_h\gamma_hE_h - (m_h + \mu_h)I_h\\\
		R_h'(t) = k_hE_h - (m_h + \lambda_h )R_h
	\end{cases}
\end{align*}
\endgroup

The flow chart of the model and the correspective parameters value are reported
in \cite{paperCINA}. The system presents both the vitality and the epidemic
dynamic. It is important to underline that the rate of loss of immunity for the
dogs population has been added, compared to the original SEIR. The basic
epidemic reproduction number can be computed \cite{23,96PapCinaBio} as:

$$
R_0=\frac{\beta_d\sigma_d\gamma_d}{(m_d+\kappa_d+\sigma_d)(m_d+\mu_d)}\frac{(m_d
+\lambda_d)A}{(m_d+\kappa_d+\lambda_d)m_d}
$$

Moreover, it can be proved that there are two equilibria:
\begin{itemize}
	\item a disease free equilibrium $E_0=(S_{d,0},0,0,R_{d,0},S_{h,0},0,0,0)$;
	\item an endemic equilibrium
	$E^*=(S^*_d,E^*_d,I^*_d,R^*_{d},S^*_{h},E^*_h,I^*_h,R^*_h)$ which exists if 
	and only if holds the condition
	\begin{equation}
		\label{c2}
		R_0>1.
	\end{equation}
\end{itemize}

The complete expression of $E^*$ as a function of the parameters can be found in
\cite{paperCINA}. If the condition \ref{c2} is true, it can be proved that $E_0$
is locally asymptotically stable in the region

\begin{align*}
  \Gamma=&\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in(\mathbb{R}_+\cup\{0\})^8~\vert\\
  &0 < S_d + E_d + I_d + R_d \le \frac{A}{m_d}\Bigr\}\\
\end{align*}

Moreover, if \ref{c2} holds, it can be shown that $E*$ is locally asymptotically
stable in $\hat{\Gamma}=\Gamma-\Sigma$, where

$$
\Sigma=\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in\Gamma~\vert~I_h=0\Bigr\}
$$
and all the solution in $\Sigma$ tend toward the disease free equilibrium. A
complete demonstration of the previous statement can be found in
\cite{113papCinaBio}.

Finally, the paramteres of the model can be estimated using the data collected
by the Chinese Ministry of Health. Using these parameters, it can be shown that
the $R_0$ value for rabies in China is around 2. Therefore the epidemic moves
towards the endemic equilibrium.

A numerical simulation of the model from 1996, for a period of 50 years, is
shown in Fig. \ref{fig:dh1}. The model indicates the tendency of the rabies
epidemic to persist and move towards the endemic equilibrium.

```{r, echo=FALSE, fig.width=7, fig.height=6,fig.cap="\\label{fig:dh1}Numerical simulations of rabies cases in China from 1996, for a period of 50 years. $S(t)$ and $R(t)$ are plotted in log$_{10}$ scale. The curve for $I(t)$ reproduces Fig. 2.2b of \\cite{original}."}
library(latex2exp)
source("SEIR_dogs_humans.R")

# ggplot(data = as.data.frame(SEIR_out)) +
#   geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
#   ylab(TeX("$I_h$")) + xlab("time (months)")

ggplot(data = SEIR_out_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab(latex2exp::TeX("$E_h$, $I_h$")) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~., name =
    latex2exp::TeX("$S_h$, $R_h$"),
    breaks = breaks, labels = labels)) + theme(legend.position="bottom",
      axis.title.y.right = element_text(margin = margin(l = 20))) + 
  xlab("time (months)")
```

We can examine how the initial conditions influence the dynamic of $I(t)$. The
initial conditions of dogs can influence both the number of human rabies cases
and the time of rabies case peak, as shown in Fig. \ref{fig:dh2} Eventually, the
differences become small. We also observe that the peak of the initial outbreak
would be postponed if $S_d(0)$ is decreasing.
 
```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dh2}Number of human rabies cases $I_h(t)$ for different initial values of the susceptible dog population $S_d(0)$. This plot reproduces Fig. 2.3a of \\cite{original}."}
ggplot(data = I_h_df_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = latex2exp::TeX("$S_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)") +
  xlim(0, 1000)
```

Instead, the initial conditions of humans influence only the number of human
rabies cases and have no effect on th time of rabies cases peak, as we can see
in Fig. \ref{fig:dh3}.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dh3}Number of human rabies cases $I_h(t)$ for different initial values of the susceptible human population $S_h(0)$. This plot reproduces Fig. 2.3b of \\cite{original}."}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = latex2exp::TeX("$S_h(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)") +
  xlim(0, 1000)
```

## SEIV model with domestic and stray dogs

Although the previous model provided a good estimation of the epidemic dynamic
of the last years, it is not completely satisfactory. As a matter of fact,
almost every region of China is populated by stray dogs. Consequently, it is
useful to introduce another model, where the parameters of stray and domestic
dogs are distinctly estimated.

The simplest formulation of this stray-domestic-human system is realized through
a SEIV model with demographic dynamic at constant coefficients. In the SEIV
model the removed class is replaced by the vaccinated class. In this case it is
assumed that only susceptible and exposed domestic flow with a specific rate in
the vaccinated class. Moreover it is assumed that domestic dogs can be abandoned
and consequently the become stray in the respective class. It is also assumed
that only domestic dogs reproduce and that the stray dogs of all classes are
culled at the same rate. For the human population it is coherently supposed that
only exposed individuals are treated, so they can move in the vaccinated class.
The model can be presented as:

\begingroup
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s4}
	\begin{cases}
		S_0'(t) = lS_1 - (\mu + c)S_0 - \beta_0S_0I_0 + p_0E_0 + \delta_1V_0\\
		E_0'(t) = lE_1 + \beta_0S_0I_0 - (\mu + c + \sigma_0 + p_0)E_0\\
		I_0'(t) = \sigma_0E_0 + \epsilon I_1 - (\mu + c + \alpha)I_0\\
		V_0'(t) = lV_1 - (\mu + c + \delta_1)V_0\\
		S_1'(t) = A - \beta S_1I_0 - \beta_1S_1I_1 - (1 + \nu + l)S_1 + p_1E_1 + \delta2V_1\\
		E_1'(t) = \beta S_1I_0 + \beta_1S_1I_1 - (1 + \nu + \sigma_1 + l + p_1)E_1\\
		I_1'(t) = \sigma_1E_1 - (1 + \epsilon + k)I_1\\
		V_1'(t) = \nu(S_1 + E_1) - (1 + l + \delta_2)V_1\\
		S_h'(t) = H - \mu_hS_h - \lambda_{1h}S_hI_0 - \lambda_{2h}S_hI_1 + \delta_{1h}Eh + \delta_{2h}V_h\\
		E_h'(t) =\lambda_{1h}S_hI_0 + \lambda_{2h}S_hI_1 - (\mu_h + \sigma_h + \delta_{1h} + \nu_h)E_h\\
		I_h'(t) = \sigma_hE_h - (\mu_h + \alpha_h )I_h\\
		V_h'(t) = \nu_hE_h - (\mu_h + \delta_{2h})V_h\\
	\end{cases}
\end{align}
\endgroup
	
It can be proved that given
\begin{align*}
  X=&\Bigl\{(S_0, E_0, I_0,V_0, S_1, E_1, I_1,V_1, S_h, E_h, I_h,V_h )\in\mathbb{R}_+^{12}~\vert\\
  &E_0, I_0, E_1, I_1, E_h, I_h,V_h\in\mathbb{R}_+\cup\{0\}\Bigr\}\\
\end{align*}
the region
\begin{align*}
\Gamma =&\Bigl\{p\in X~\vert~S_0 + E_0 + I_0 + V_0 + S_1 + E_1 + I_1 +\\
& V_1 + S_h + E_h +I_h+V_h\le\frac{A+H}{\min\{\mu + c, d,\mu_h\}}\Bigr\}\\
\end{align*}
is positively invariant.

It is proved that there are two equilibria 
\begin{itemize}
	\item a disease free equilibrium
	$E_0=(S_{0,0},0,0,V_{0,0},S_{1,0},0,0,$ $V_{1,0},S_{h,0},0,0,0)$ where the
	complete expression is reported in \cite{paperNostro};
	\item an endemic equilibrium $E*$ whose expression is  to complicate to
	formally compute, which can be proved to exist if $R_0>1$\cite{papernostro}.
\end{itemize}

After having computed $R_0$, whose expression is presented in
\cite{paperNostro}, it can be proved that if $R_0<1$, $E_0$ is globally
asymptotically stable for system \ref{s4}.

Numerical simulations can be performed using paramteres estimated from data
provided by the Department of Health of Guangdong Province. Fig. \ref{fig:dsh1}
and Fig. \ref{fig:dsh2} show a simulation of human rabies infection cases over
time for the Guangdong Province.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}Simulation of human rabies infection cases from 2006 to 2014 for the Guangdong Province of China. This plot reproduces Fig. 3.2 of \\cite{original}."}
source("SEIR_dogs_stray_humans.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme +
  ylab(TeX("$I_h$")) + xlab("time (months)") + xlim(0, 100)
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh2}Numerical simulations of rabies cases in the Guangdong Province from 2006 to 2014. $S(t)$ and $V(t)$ are plotted in log$_{10}$ scale."}
ggplot(data = SEIR_out_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab(latex2exp::TeX("$E_h$, $I_h$")) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~., name =
      latex2exp::TeX("$S_h$, $V_h$"),
    breaks = breaks, labels = labels)) + theme(legend.position="bottom",
      axis.title.y.right = element_text(margin = margin(l = 20))) + 
  xlab("time (months)")
```

We can examine how different values of the parameters influence the dynamic of
$I(t)$. Fig. \ref{fig:dsh3} shows the dynamic of $I_h(t)$ for different values
of $l$, the rate of domestic dog abandonment, and $\epsilon$, the rate of
infectious domestic dog abandonment. We can observe that reducing $\epsilon$ is
more effective than reducing $l$.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh3}Simulations of infected human rabies cases $I_h(t)$ in the Guangdong Province for different values of $l$ and $\\epsilon$. This plot reproduces Fig. 3.5a of \\cite{original}."}
ggplot(data = I_h_df_long_1, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$l$, $\\epsilon$"), labels = labels_1) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)") + xlim(0, 125)
```

Fig. \ref{fig:dsh4} shows the dynamic of $I_h(t)$ for different values of $c$,
the culling rate of stray dogs, and $\nu$, the vaccination rate of domestic
dogs. We can observe that an increasing the vaccination rate is more effective
in reducing $I_h(t)$ than increasing the culling rate.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh4}Simulations of infected human rabies cases $I_h(t)$ in the Guangdong Province for different values of $c$ and $\\nu$. This plot reproduces Fig. 3.5b of \\cite{original}."}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "", labels = c(
    "No intervention", "50% stray dog culling", "100% stray dog culling",
    "50% vaccination", "70% vaccination")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + guides(col = guide_legend(nrow = 3)) + xlim(0, 125)
```

Finally, Fig. \ref{fig:dsh5} shows the dynamic of $I_h(t)$ for different values
of $\nu_h$, the culling rate of stray dogs, the vaccination rate of exposed
humans. Increasing $\nu_h$ decreases $I_h(t)$, but does not stop the epidemic.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh5}Simulations of infected human rabies cases $I_h(t)$ in the Guangdong Province for different values of $\\nu_h$. This plot should reproduces Fig. 3.5c of \\cite{original}, but is quite different."}
ggplot(data = I_h_df_long_3, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$\\nu_h$"), labels = nu_h_vect) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + xlim(0, 125)
```

## Seasonal SEIR model

It is quite evident that the assumption of a constant recruitment rate it is not
biologically plausible \cite{54biopapnostro}. Moreover it has been observed that
also the contact rate and vaccination rate depends on time
\cite{78,biopapnostro}. Therefore there is a periodic component in the rabies
epidemic phenomenon that the previous model do not grasp. A new model has been
presented to highlight the periodicity component. It is a SEIR model with
periodic transmission rate:
\begingroup
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s5}
	\begin{cases}
		S_d'(t) = A + \lambda_dR_d + \sigma_d (1 - \gamma_d )E_d - \beta_d(t) S_dI_d - (m_d + k_d)S_d\\
		E_d'(t) = \beta_d(t) S_dI_d - \sigma_d (1 - \gamma_d )E_d - \sigma_d\gamma_d E_d - (m_d + k_d)E_d\\
		I_d'(t) = \sigma_d\gamma_d E_d - (m_d + \mu_d)I_d\\
		R_d'(t) = k(S_d + E_d) - (m_d + \lambda_d)R_d\\
		S_h'(t) = B + \lambda_hR_h + \sigma_h(1 - \gamma_h )E_h - m_hS_h - \beta_{dh}(t)S_hI_h\\
		E_h'(t) = \beta_{dh}(t)S_hI_d - \sigma_h(1 - \gamma_h )E_h - \sigma_h\gamma_hE_h - (m_h + k_h )E_h\\
		I_h'(t) = \sigma_h\gamma_hE_h - (m_h + \mu_h)I_h\\\
		R_h'(t) = k_hE_h - (m_h + \lambda_h )R_h
	\end{cases}
\end{align}
\endgroup

where $\beta_d(t)=a[1 + bsin( \frac{\pi}{6} t + 5.5)]$ and
$\beta_{dh}(t)=a_1[1 + b_1sin( \frac{\pi}{6} t + 5.5)]$. Even in this case, set 
$$
X=\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in (\mathbb{R}_+\cup\{0\})^8\Bigr\}
$$

it can be proved that there is a positively invariant region
\begin{align*}
\Gamma =&\Bigl\{p\in X~\vert~0 < S_d + E_d + I_d + R_d \le \frac{A}{m_d}\text{ and }\\
&0 < S_h + E_h + I_h + R_h \le \frac{B}{m_h}\Bigr\}\\
\end{align*}
It is proved that the system \ref{s5} has 
\begin{itemize}
	\item a disease free equilibrium $E_0=(S_{0},0,0,R_{0},S_{h},0,0,0)$ where the
	complete expression is reported in \cite{paperNostro}, which is globally
	asymptotically stable for $R_0<1$;
	\item at least one positive periodic solution \cite{111paperNostroBio}.
\end{itemize} 

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s1}plotting example"}
source("SEIR_seasonal.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
  ylab(TeX("$I_h$")) + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s2}plotting example"}
ggplot(data = SEIR_out_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "class") +
  myTheme + ylab(latex2exp::TeX("$E_h$, $I_h$")) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~., name =
      latex2exp::TeX("$S_h$, $R_h$"),
    breaks = breaks, labels = labels)) + theme(legend.position="bottom",
      axis.title.y.right = element_text(margin = margin(l = 20))) + 
  xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s3}plotting example"}
ggplot(data = as.data.frame(SEIR_out_2)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
  ylab(TeX("$I_h$")) + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s4}plotting example"}
ggplot(data = I_h_df_long_1, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$S_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + guides(col = guide_legend(nrow = 2))
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:s5}plotting example"}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$I_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)")
```


## Multi-patch SEIV model

In 2005 the Chinese government decided to monitor rabies at the national level
in an attempt to obtain a more comprehensive epidemiological dataset. The data
highlight that east China is epidemiologically related to adjoining provinces
and to distant provinces. Moreover it has been demonstrated that there is a
general trend of longitudinal transmission and latitudinal transmission,
\cite{106}, \cite{108}.
	
A new SIEV model is developed to analyze how the movements of dogs affects the
spread of rabies. In order to include this aspect, it has not been used a
reaction-diffusion model because it is not biological coherent to suppose
continuous the movement of the dogs and connected the spatial domain. It is
preferred to divide the entire China region in smaller area, called patch. Then
two travel rate matrix $\Phi$ and $\Psi$ are estimated such that each class of
dogs or humans respectively move from a patch to another. The obtained model is:
\begingroup
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s6}
	\begin{cases}
	S_{d,i}'(t) = & A_i + \lambda_{d,i}V_{d,i} + \sigma_{d,i} (1 - \gamma_{d,i})E_{d,i} - \beta_{d,i}S_{d,i}I_{d,i} +\\
	& - (m_{d,i} + k_{d,i})S_{d,i}+\sum_{j=1}^n\Phi^S_{ij}S_{d,j}\\
	E_{d,i}'(t) = & \beta_{d,i}  S_{d,i}I_{d,i} - \sigma_{d,i} (1 - \gamma_{d,i}
	)E_{d,i} - \sigma_{d,i}\gamma_{d,i} E_{d,i} +\\
	& - (m_{d,i} + k_{d,i})E_{d,i}+\sum_{j=1}^n\Phi^E_{ij}E_{d,j}\\
	I_{d,i}'(t) = & \sigma_{d,i}\gamma_{d,i} E_{d,i} - (m_{d,i} +
	\mu_{d,i})I_{d,i}+\sum_{j=1}^n\Phi^I_{ij}I_{d,j}\\
	V_{d,i}'(t) =& k_{d,i}(S_{d,i} + E_{d,i}) - (m_{d,i} + \lambda_{d,i})V_{d,i}
	+\sum_{j=1}^n\Phi^V_{ij}V_{d,j}\\
	S_{h,i}'(t) =& B + \lambda_{h,i}V_{h,i} + \sigma_{h,i}(1 - \gamma_{h,i} )E_{h,i}
	- m_{h,i}S_{h,i} +\\
	& - \beta_{dh,i}S_{h,i}I_{h,i}+\sum_{j=1}^n\Psi^S_{ij}S_{h,j}\\
	E_{h,i}'(t) =& \beta_{dh,i} S_{h,i}I_{d,i} - \sigma_{h,i}(1 - \gamma_{h,i}
	)E_{h,i} - \sigma_{h,i}\gamma_{h,i}E_{h,i} +\\
	& - (m_{h,i} + k_{h,i} )E_{h,i}	+\sum_{j=1}^n\Psi^E_{ij}E_{h,j}\\
	I_{h,i}'(t) =& \sigma_{h,i}\gamma_{h,i}E_{h,i} - (m_{h,i} + \mu_{h,i})I_{h,i} +
	\sum_{j=1}^n\Psi^I_{ij}I_{h,j}\\
	V_{h,i}'(t) =& k_{h,i}E_{h,i} - (m_{h,i} + \lambda_{h,i}
	)V_{h,i}+\sum_{j=1}^n\Psi^V_{ij}V_{h,j}\\
	\end{cases}
\end{align}
\endgroup

where  with subscript $d$ we denote dogs class and with subscript $h$ human
ones. It is a classical SEIV model for every patch $i$ for $i\in\{1,\dots,n\}$.
It is important to underline that
$$
\Phi^X_{ii}=-\sum_{\substack{j=1 \\ j\neq i}}^n\Phi^X_{ij} \quad\text{and}\quad\Psi^X_{ii}=-\sum_{\substack{j=1 \\ j\neq i}}^n\Psi^X_{ij} 
$$
for $i\in\{1,\dots,n\}$ and for $X\in\{S,E,I,V\}$. Moreover, $\Phi^X_{ij}\ge0$
and $\Psi^X_{ij}\ge0$ for $i,j\in\{1,\dots,n\}$ and $X\in\{S,E,I,V\}$.

It can be proved that there exist a disease free equilibrium
$E_0=(\underline{S}_{d,0},\underline{0},\underline{0},\underline{V}_{d,0},\underline{S}_{h,0},\underline{0},\underline{0},\underline{0})$, whose expression is presented in \cite{Cina}.

Even trying to compute the formal expression of $R_0$ is extremely complicated
for two patches. Therefore the the study of this model behaviuor is based on the
numerical simulations.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:p1}plotting example"}
source("SEIV_patch.R")

ggplot(data = out, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "Province", labels =
      c("Hebei", "Fujian")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + xlim(0, 100)
```

# Novel results
MC : eq, plots, flow



## Multi-patch SEIV model with stray dogs

```{r, include=FALSE}
wdgt <- DiagrammeR::grViz("patch_SEIV.gv")
wdgt_html_tf <- tempfile(fileext=".html")
htmlwidgets::saveWidget(wdgt, wdgt_html_tf)
webshot::webshot(url=sprintf("file://%s", wdgt_html_tf), 
  selector="#htmlwidget_container",
  file="patch_SEIV.pdf")
unlink(wdgt_html_tf)
```
\begin{figure*}[!htb]
  \centering
  \includegraphics[width=\textwidth,trim = 4cm 7.1cm 2.5cm 1.2cm, clip]{patch_SEIV.pdf}
  \caption{test}
  \label{fig:mess}
\end{figure*}

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:ps1}plotting example"}
source("SEIV_patch_stray.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h_1), color = "blue") + 
  geom_line(mapping = aes(time, I_h_2), color = "red") + myTheme
```

## Multi-patch seasonal SEIV model

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:pss1}plotting example"}
source("SEIV_patch_stray_seasonal.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h_1), color = "blue") + 
  geom_line(mapping = aes(time, I_h_2), color = "red") + myTheme
```

# Conclusions

Different models have been proposed and summarized to describe the dynamic of the rabies epidemic in China. However some small inconsistencies in \cite{Cina} has made studying this paper challenging. Moreover data deficit has made impossible to estimate the accuracy of the new proposed patches model.

\begin{thebibliography}{1}
	
	\bibitem{Wang}Wang L, Tang Q, Liang G. Rabies and rabies virus in wildlife in
	mainland China, 1990-2013. Int J Infect Dis. 2014; 25:122-9.
	
	\bibitem{Schnell}Schnell MJ, McGettigan JP, Wirblich C, Papaneri A. The cell
	biology of rabies virus: using stealth to reach the brain. Nat Rev Microbiol.
	2010; 8(1):51-61.
	
	\bibitem{Velasco}Velasco-Villa A, Escobar LE, Sanchez A, Shi M, Streicker DG,
	Gallardo-Romero NF, Vargas-Pino F, Gutierrez-Cedillo V, Damon I, Emerson G.
	Successful strategies implemented towards the elimination of canine rabies in
	the Western Hemisphere. Antiviral Res. 2017; 143:1-12.
	
	\bibitem{Hemachudha}Hemachudha T, Ugolini G, Wacharapluesadee S, Sungkarat W,
	Shuangshoti S, Laothamatas J. Human rabies: neuropathogenesis, diagnosis, and
	management. Lancet Neurol. 2013; 12(5):498-513.
	
	\bibitem{Cina} 
	Ruan S.Modeling the transmission dynamics and control of rabies in China.
	Mathematical biosciences 286. 2017; 65-93.
	
	\bibitem{54} 
	J. Ma , Z. Ma. Epidemic threshold conditions for seasonally forced SEIR models.
	Math. Biosci. Eng. 3. 2006; 161 -172.
	 
	\bibitem{78} 
	I. Schwartz. Small amplitude, long periodic out breaks in seasonally driven
	epidemics. J. Math. Biol. 30. 1992; 473 -491 .
	
	\bibitem{113} 
	J. Zhang, Z. Jin, G. Sun, T. Zhou, S. Ruan. Analysis of rabies in China:
	transmission dynamics and control. PLoS One 6. 2011; e20891.
	
	\bibitem{111} 
	J. Zhang, Z. Jin, G. Sun, T. Zhou, S. Ruan. Analysis of rabies in China:
	transmission dynamics and control. PLoS One 6. 2011; e20891.
	
	\bibitem{108} 
	J. Yu, L. Hao, Q. Tang. The spatial and temporal dynamics of rabies in China.
	PLoS Negl. Trop. Dis. 6. 2012; e1640.
	
	\bibitem{106} 
	C. Yin, H. Zhou, H. Wu. Analysis on factors related to rabies epidemic in China
	from 2007 -2011. Virol. Sin. 27. 2012; 132 -143.
	
	\bibitem{23} 
	Diekmann, J.A.P. Heesterbeek, M.G. Roberts. The construction of next-generation
	matrices for compartmental epidemic models. J. R. Soc. Interface 7. 2010; 873
	-885.
	
	\bibitem{96} 
	X. Wang, J. Lou. Two dynamic models about rabies between dogs and human. Biol.
	Syst. 16. 2008; 519 -529.
	
	\bibitem{mate1} 
	Stefan Ma, Yingcun Xia. Mathematical Understanding of Infectious Disease
	Dynamics. Lecture Notes Series, Institute for Mathematical Sciences, National
	University of Singapore: Volume 16. 2008; 33-35.

\end{thebibliography}

