---
title: "Rabies"
date: "`r Sys.Date()`"
author:
-   name: Matteo Ciciani
    affiliation: &aff Centre for Integrative Biology (CIBIO), University of
        Trento, Italy
-   name: Martina Iannacito
    affiliation: *aff
output:
    BiocStyle::pdf_document:
      fig_caption: yes
vignette: >
    %\VignetteIndexEntry{Signature-based Clustering for Diagnostic Purposes}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

## Rabies

Rabies is one of the oldest diseases in recorded human history, dating back more
than 4000 years. More than 70000 human rabies deaths occur annually, and Asia
accounts for approximately 80\% of the worldwide total \cite{Wang}. As 40\% of
those infected are children, the years of life lost by this infection makes
rabies the seventh most important global infectious disease \cite{Schnell}.
Currently, China is second only to India in the number of people killed by
rabies worldwide, and the epidemic area has expanded to almost the entire
country \cite{Wang}.

Rabies is an acute encephalomyelitis (inflammation of the brain and the spinal
cord) caused by viruses in the genus \emph{Lyssavirus} of the family
\emph{Rhabdoviridae}. Rabies lyssavirus (RABLV), the prototype species of the
genus, causes disease in nearly all terrestrial mammals. Despite the worldwide
distribution of RABLV, only a few animal species in the orders \emph{Carnivora},
\emph{Chiroptera} (bats) and \emph{Primates} play a critical role in maintaining
the circulation of the virus in nature. Most of these mammals harbor
species-specific RABLV variants which circulate within discrete geographic
boundaries. However, because almost all cases of human rabies result from the
bites of rabid dogs, dog-maintained viruses are by far the most important target
of control measures \cite{Velasco}.

As for most viruses, the rabies virus life cycle can be divided into three
phases. The first phase includes entry into the host cell by endocytosis
followed by release of the viral genome into the cytoplasm. The classical
wild-type rabies virus first infects a muscle cell, then enters a motor neuron
at the neuromuscular junction. Rabies virus particles are subsequently
transported in a retrograde direction through the axon of the infected neuron.
Once the virus particles reach the cell body of the neuron, the second phase
commences, which involves the production of the virion components. The last
phase of the life cycle comprises the assembly of the viral components and the
release of mature rabies virus particles, which can then start a new round of
infection \cite{Schnell}.

By the time a patient develops the first prodromal symptoms, such as fever,
flu-like symptoms, and gastrointestinal disturbances, the virus is already
widely disseminated throughout the central nervous system \cite{Hemachudha}.

Clinical manifestation of rabies in humans has two forms: the furious form (80\%
of infections), and the paralytic form (20\% of infections). The furious form of
rabies is characterized by hydrophobia, episodes of hallucinations and
excitement. These behavioral changes occur simultaneously with the shedding of
large amounts of rabies virus in the saliva, which facilitates the spread of the
virus to a new host. The paralytic form of rabies is characterized by weakness
and flaccid paralysis. In both cases, survival after the onset of symptoms is
rarely more than 7 days and the disease is lethal in almost 100\% of cases
\cite{Schnell}.

Human rabies virus infections are controlled by limiting the infection in
animals and by post-exposure prophylaxis (PEP) of humans after contact with a
potentially infected animal. Current PEP for previously non-immunized
individuals consists of one dose with rabies immunoglobulin in conjunction with
five doses of immunization with inactivated rabies virus vaccines. This regimen
is effective, but its usefulness in developing countries, where rabies is
endemic, is hampered by the high cost and lack of compliance \cite{Schnell}.

## SEIR

Nowadays mathematics is used in different scientific areas. As a matter of fact
mathematical models are commonly applied to investigate the evolution of
epidemics. One of the most known model is SEIR. It is coherent to apply a basic
SEIR model if the considered disease has an average period shorter than the
average time needed to observe significant modifications in the population
structure. For example it can be applied to describe flu behaviour into the
human specie. As a matter of fact life of a human being and generation of a new
individ- ual takes on average more than the duration of the flu. Under this
assumption demographical aspects are negligible and consequently the model focus
only on the disease dynamic. In order to apply the model the total population
$N(t)$ at time t is split into four class:

- $S(t)$ : it is the class of susceptible to the disease at time t;
- $E(t)$ : it is the class of exposed to the pathogen and not contagious at time
t;
- $I(t)$ : it is the class of infectuous at time t;
- $R(t)$ : it is the class of removed, i.e. those who get out of the model
because they are dead or they have become immune at time t.

$$
\begin{cases}
x = 1+1\\
y = 3
\end{cases}
$$

Then it is assumed that susceptible ones who come into contact with infected
become exposed at rate $\lambda(t)$ , exposed move into the infectuous class at rate
$k(t)$ and at last infectous are removed at rate $\gamma(t)$ . The force of
infection $\lambda(t)$ is mathematically described as:

$$
\lambda(t) = c(t)\chi\frac{I(t)}{N(t)}
$$

where $c(t)$ is the average number of contact of one individual in the unit of
time, $\chi$ is the probability to become infect after a contact with $I(t)$ an
infectuous and $\frac{I(t)}{N(t)}$ is the total number of infectuous ones. To
make computations easier it is assumed that all these parameter are constant in
time and thanks to the absence of demographic aspects also the total population
is constant. Therefore the SEIR ode system is:

\begin{align}
  \label{1}
  \begin{cases}
    x = 1+1\\
    y = 3
  \end{cases}
\end{align}

It is important to underline that if $\gamma$ is the rate at which an infectuous
is removed, then $\gamma$ 1 is the average time an individual one spends as
infectuous. Consequently the basic epidemiologic reproduction number is the
number of ill people generated by an infectuous in the the life time as
infectuous, i.e.

$$
R_0 = \frac{c\chi}{\gamma}
$$

Add other systems

Apply SEIR to rabies in China

# Reproduction of original results


## Dog + humans 

MI : equations, stability

```{r figs, echo=FALSE, fig.width=7,fig.height=6,fig.cap="\\label{fig:figs}plotting example"}
source("SIER_dogs_humans.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme

```

## Stray +  Domestic + human

MI : equations

MC : plots

## Seasonal

MI : equation
MC : plots

## Patch

MI: eq
MC: plot

# New stuff

MC : eq, plots, flow

## Patch + stray

## Patch + stray + seasonal

# Conclusion

MI

\begin{thebibliography}{1}
	
	%Formato PLOS	
	
	\bibitem{Wang}Wang L, Tang Q, Liang G. Rabies and rabies virus in wildlife in mainland China, 1990-2013. Int J Infect Dis. 2014; 25:122-9.
	
	\bibitem{Schnell}Schnell MJ, McGettigan JP, Wirblich C, Papaneri A. The cell biology of rabies virus: using stealth to reach the brain. Nat Rev Microbiol. 2010; 8(1):51-61.
	
	\bibitem{Velasco}Velasco-Villa A, Escobar LE, Sanchez A, Shi M, Streicker DG, Gallardo-Romero NF, Vargas-Pino F, Gutierrez-Cedillo V, Damon I, Emerson G. Successful strategies implemented towards the elimination of canine rabies in the Western Hemisphere. Antiviral Res. 2017; 143:1-12.
	
	\bibitem{Hemachudha}Hemachudha T, Ugolini G, Wacharapluesadee S, Sungkarat W, Shuangshoti S, Laothamatas J. Human rabies: neuropathogenesis, diagnosis, and management. Lancet Neurol. 2013; 12(5):498-513.
	
\end{thebibliography}
