---
title: SIER model for rabies in China
author: Martina Iannacito\textsuperscript{1} and Matteo Ciciani\textsuperscript{2}
affiliation:
  - \textsuperscript{1}\textit{...}
  - \textsuperscript{2}\textit{Quantitative and Computational Biology Master, University of Trento, Italy}
output:
    pdf_document:
      latex_engine: xelatex
      template: SelfArx-template.tex
abstract: Here is the abstract
---
# Introduction

## The biology of rabies
Rabies is one of the oldest diseases in recorded human history, dating back more
than 4000 years. More than 70000 human rabies deaths occur annually, and Asia
accounts for approximately 80\% of the worldwide total \cite{Wang}. As 40\% of
those infected are children, the years of life lost by this infection makes
rabies the seventh most important global infectious disease \cite{Schnell}.
Currently, China is second only to India in the number of people killed by
rabies worldwide, and the epidemic area has expanded to almost the entire
country \cite{Wang}.

Rabies is an acute encephalomyelitis (inflammation of the brain and the spinal
cord) caused by viruses in the genus \emph{Lyssavirus} of the family
\emph{Rhabdoviridae}. Rabies lyssavirus (RABLV), the prototype species of the
genus, causes disease in nearly all terrestrial mammals. Despite the worldwide
distribution of RABLV, only a few animal species in the orders \emph{Carnivora},
\emph{Chiroptera} (bats) and \emph{Primates} play a critical role in maintaining
the circulation of the virus in nature. Most of these mammals harbor
species-specific RABLV variants which circulate within discrete geographic
boundaries. However, because almost all cases of human rabies result from the
bites of rabid dogs, dog-maintained viruses are by far the most important target
of control measures \cite{Velasco}.

As for most viruses, the rabies virus life cycle can be divided into three
phases. The first phase includes entry into the host cell by endocytosis
followed by release of the viral genome into the cytoplasm. The classical
wild-type rabies virus first infects a muscle cell, then enters a motor neuron
at the neuromuscular junction. Rabies virus particles are subsequently
transported in a retrograde direction through the axon of the infected neuron.
Once the virus particles reach the cell body of the neuron, the second phase
commences, which involves the production of the virion components. The last
phase of the life cycle comprises the assembly of the viral components and the
release of mature rabies virus particles, which can then start a new round of
infection \cite{Schnell}.

By the time a patient develops the first prodromal symptoms, such as fever,
flu-like symptoms, and gastrointestinal disturbances, the virus is already
widely disseminated throughout the central nervous system \cite{Hemachudha}.

Clinical manifestation of rabies in humans has two forms: the furious form (80\%
of infections), and the paralytic form (20\% of infections). The furious form of
rabies is characterized by hydrophobia, episodes of hallucinations and
excitement. These behavioral changes occur simultaneously with the shedding of
large amounts of rabies virus in the saliva, which facilitates the spread of the
virus to a new host. The paralytic form of rabies is characterized by weakness
and flaccid paralysis. In both cases, survival after the onset of symptoms is
rarely more than 7 days and the disease is lethal in almost 100\% of cases
\cite{Schnell}.

Human rabies virus infections are controlled by limiting the infection in
animals and by post-exposure prophylaxis (PEP) of humans after contact with a
potentially infected animal. Current PEP for previously non-immunized
individuals consists of one dose with rabies immunoglobulin in conjunction with
five doses of immunization with inactivated rabies virus vaccines. This regimen
is effective, but its usefulness in developing countries, where rabies is
endemic, is hampered by the high cost and lack of compliance \cite{Schnell}.

## SEIR models

Nowadays mathematics is used in different scientific areas. As a matter of fact
mathematical models are commonly applied to investigate the evolution of
epidemics. One of the most known model is SEIR. It is coherent to apply a basic
SEIR model if the considered disease has an average period shorter than the
average time needed to observe significant modifications in the population
structure. For example it can be applied to describe flu behaviour into the
human specie. As a matter of fact life of a human being and generation of a new
individ- ual takes on average more than the duration of the flu. Under this
assumption demographical aspects are negligible and consequently the model focus
only on the disease dynamic. In order to apply the model the total population
$N(t)$ at time t is split into four class:
\begin{itemize}
  \item $S(t)$ : it is the class of susceptible to the disease at time t;
  \item $E(t)$ : it is the class of exposed to the pathogen and not contagious
  at time t;
  \item $I(t)$ : it is the class of infectuous at time t;
  \item $R(t)$ : it is the class of removed, i.e. those who get out of the model
  because they are dead or they have become immune at time t.
\end{itemize}

Then it is assumed that susceptible ones who come into contact with infected
become exposed at rate $\lambda(t)$ , exposed move into the infectuous class at
rate $k(t)$ and at last infectous are removed at rate $\gamma(t)$ . The force of
infection $\lambda(t)$ is mathematically described as:

$$
\lambda(t) = c(t)\chi\frac{I(t)}{N(t)}
$$

where $c(t)$ is the average number of contact of one individual in the unit of
time, $\chi$ is the probability to become infect after a contact with $I(t)$ an
infectuous and $\frac{I(t)}{N(t)}$ is the total number of infectuous ones. To
make computations easier it is assumed that all these parameter are constant in
time and thanks to the absence of demographic aspects also the total population
is constant. Therefore the SEIR ode system is:

\begin{align}
  \label{s1}
  \begin{cases}
    S'(t) = -\frac{c\chi}{N}S(t)I(t)\\
    E'(t) = \frac{c\chi}{N}S(t)I(t)-kE(t)\\
    I'(t) = kE(t)-\gamma I(t)\\
    R'(t) =\gamma I(t)
  \end{cases}
\end{align}

It is important to underline that if $\gamma$ is the rate at which an infectuous
is removed, then $\gamma$ 1 is the average time an individual one spends as
infectuous. Consequently the basic epidemiologic reproduction number is the
number of ill people generated by an infectuous in the the life time as
infectuous, i.e.

$$
R_0 = \frac{c\chi}{\gamma}
$$

It is possible to obtain a simpler system rescaling \ref{s1} in the following
way $\tau = \gamma t$, $u(t) =\frac{S(t)}{N}$, $v(t) = \frac{E(t)}{N}$, $w(t) =
\frac{I(t)}{N}$ and $z(t) =\frac{R(t)}{N}$. Thanks to the applied rescalation,
it follows that $u+v+w+z=\frac{S+E+I+R}{N}=1$. Denoting
$\dot{u}=\frac{du}{d\tau}$, the system (\ref{s1}) becomes:
\begin{align*}
	\begin{cases}
		\dot{u}(\tau) =-\frac{c\chi}{\gamma}u(\tau)w(\tau)\\
		\dot{v}(\tau) = \frac{c\chi}{\gamma}u(\tau)w(\tau)-\rho v(\tau)\\
		\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
		\dot{z}(\tau) = w(\tau)
	\end{cases}
\end{align*}

which is equivalent to
\begin{align}
	\label{s2}
	\begin{cases}
		\dot{u}(\tau) = -R_0u(\tau)u(\tau)w(\tau)\\
		\dot{v}(\tau) = R_0u(\tau)w(\tau)-\rho v(\tau)\\
		\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
		\dot{z}(\tau) = w(\tau).
	\end{cases}
\end{align}

It can be proved as in \cite{mate1} that given positive initial condition for
the system (\ref{s2}) there is a unique positive solution.\\ In order to make
more realistic the model it is convenient to introduce a simple vital dynamic
with $\mu$ death rate equal to $\beta$ birth rate. Under this hypothesis the
population is still constant in time. The produced system is:
\begin{align}
  \label{s3}
	\begin{cases}
    \dot{u}(\tau) = -R_0u(\tau)u(\tau)w(\tau)+\beta(1-u(\tau))\\
    \dot{v}(\tau) = R_0u(\tau)w(\tau)-(\rho+\beta)v(\tau)\\
    \dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
    \dot{z}(\tau) = w(\tau)-\beta z.
  \end{cases}
\end{align}

Reminding that $u+v+z+w=1$, it is easy to compute the equilibria of the system
(\ref{s3}), that are:
\begin{itemize}
	\item the disease free equilibrium $E_0=(1,0,0,0)$;
	\item the endemic equilibrium $E^*=(u^*,v^*,w^*,z^*)$, that exists if and only
	if holds the condition
	\begin{equation}
  	\label{c1}
  	R_0>1+\frac{\beta}{\rho}
	\end{equation}
\end{itemize}

To investigate the stability of the equilibria, it is useful to consider the
jacobian of the reduced system:
\begin{align*}
	\begin{cases}
  	\dot{u}(\tau)=-R_0u(\tau)u(\tau)w(\tau)+\beta(1-u(\tau))\\
  	\dot{v}(\tau) = R_0u(\tau)w(\tau)-(\rho+\beta)v(\tau)\\
  	\dot{w}(\tau) = \rho v(\tau)-w(\tau)\\
	\end{cases}
\end{align*}

For the disease free equilibrium the Jacobian is
$$
J(E_0)=\begin{bmatrix} 
-\beta & 0  & -R_0 \\ 
0 &	-(\rho+\beta)	& R_0\\
0 &	\rho	& -1\\
\end{bmatrix}
$$

It can be observed that it is a block matrix. Consequently an eigenvalue is
$\lambda_1=-\beta$, the other ones are the eigenvalues of the matrix
$$
J(E_0)=\begin{bmatrix} 
-(\rho+\beta)	& R_0\\
\rho	& -1\\
\end{bmatrix}
$$
which has negative trace and positive determinant it the condition (\ref{c1})
is false. For the endemic equilibrium the Jacobian is
$$
J(E^*)=\begin{bmatrix} 
-(R_0w^*+\beta) & 0  & -R_0u^* \\ 
R_0w^* &	-(\rho+\beta)	& R_0u^*\\
0 &	\rho	& -1\\
\end{bmatrix}.
$$
Using the Routh-Hurwitz conditions, it can be proved as in \cite{mate1} that the
endemic equilibrium, when it exists, is asymptotically stable.

The basic structure of the SEIR system can be adequately modified and used to
describe the rabies epidemic behavior as in \cite{PAPERNOSTRO?}.

# Reproduction of original results


## Dog + humans 

As it has been pointed out in the biological introduction to rabies, after the
bite of an infect dog, the virus has an incubation period. Consequently the
bitten human can be considered exposed. It is therefore appropriate to describe
this epidemic phenomena through a SEIR model for the dogs and the human
population. The most important difference with the classical SEIR is that in
this model only the exposed which are subjected to prophylactic measures moves
into the removed class. Besides the removed ones can loose immunity and get back
to susceptible class.

\begingroup
\setlength{\mathindent}{0pt}
\begin{align*}
	\begin{cases}
		S_d'(t) = A + \lambda_dR_d + \sigma_d (1 - \gamma_d )E_d - \beta_d S_dI_d - (m_d + k_d)S_d\\
		E_d'(t) = \beta_d S_dI_d - \sigma_d (1 - \gamma_d )E_d - \sigma_d\gamma_d E_d - (m_d + k_d)E_d\\
		I_d'(t) = \sigma_d\gamma_d E_d - (m_d + \mu_d)I_d\\
		R_d'(t) = k(S_d + E_d) - (m_d + \lambda_d)R_d\\
		S_h'(t) = B + \lambda_hR_h + \sigma_h(1 - \gamma_h )E_h - m_hS_h - \beta_{dh}S_hI_h\\
		E_h'(t) = \beta_{dh}S_hI_d - \sigma_h(1 - \gamma_h )E_h - \sigma_h\gamma_hE_h - (m_h + k_h )E_h\\
		I_h'(t) = \sigma_h\gamma_hE_h - (m_h + \mu_h)I_h\\\
		R_h'(t) = k_hE_h - (m_h + \lambda_h )R_h
	\end{cases}
\end{align*}
\endgroup

The flow chart of the model and the correspective parameters value are reported
in \cite{paperCINA}. The system presents both the vitality and the epidemic
dynamic. It is important to underline that it has been added with respet to the
original SEIR the rate of loss of immunity for the dogs population. It can be
computed, as in \cite{23,96PapCinaBio} the basic epidemic reproduction number
as:
$$
R_0=\frac{\beta_d\sigma_d\gamma_d}{(m_d+\kappa_d+\sigma_d)(m_d+\mu_d)}\frac{(m_d
+\lambda_d)A}{(m_d+\kappa_d+\lambda_d)m_d}
$$

Moreover it can be proved that there are two equilibria:
\begin{itemize}
	\item a disease free equilibrium $E_0=(S_{d,0},0,0,R_{d,0},S_{h,0},0,0,0)$;
	\item an endemic equilibrium
	$E^*=(S^*_d,E^*_d,I^*_d,R^*_{d},S^*_{h},E^*_h,I^*_h,R^*_h)$ which exists if 
	and only if holds the condition
	\begin{equation}
		\label{c2}
		R_0>1.
	\end{equation}
\end{itemize}

The complete expression of $E^*$ in function of the parameters can be found in
\cite{paperCINA}. If the condition \ref{c2} is it can be proved that $E_0$ is
locally asymptotically stable in the region

\begin{align*}
  \Gamma=&\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in(\mathbb{R}_+\cup\{0\})^8~\vert\\
  &0 < S_d + E_d + I_d + R_d \le \frac{A}{m_d}\Bigr\}\\
\end{align*}

Moreover if \ref{c2} holds, it can be shown that $E*$ is locally asymptotically
stable in $\hat{\Gamma}=\Gamma-\Sigma$, where

$$
\Sigma=\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in\Gamma~\vert~I_h=0\Bigr\}
$$
and all the solution in $\Sigma$ tend toward the disease free equilibrium. A
complete demonstration of the previous statement can be found in
\cite{113papCinaBio}.\\ At last a simulation of the model can be realized using
the data collected by the Chinese Ministry of Health. After a complete analysis
was performed, the researchers could show that the $R_0$ value in China is
around $1.2$. Therefore the epidemic moves toward the endemic equilibrium. But
the analysis was able to suggest possible ways of controlling the epidemic,
through reduction of the annual crop of puppies and increasing the dog
immunization rate.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dh1}plotting example"}
library(latex2exp)
source("SEIR_dogs_humans.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
  ylab(TeX("$I_h$")) + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dh1}plotting example"}
ggplot(data = I_h_df_long, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = latex2exp::TeX("$S_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dh1}plotting example"}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = latex2exp::TeX("$S_h(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)")
```

## Stray +  Domestic + human

Although the previous model provided a good estimation of the Chinese rabies
epidemic dynamic of the last years, it is not completely satisfactory. As a
matter of fact almost every region of China is populated by stray dogs.
Consequently it is useful to introduce another model, where the parameters of
stray and domestic dogs are distinctly estimated.

The simplest formulation of this stray-domestic-human system is realized through
a SEIV with demographic dynamic at constant coefficients. In the SEIV model the
removed class is replaced by the vaccinated class. In this case it is assumed
that only susceptible and exposed domestic flow with a specific rate in the
vaccinated class. Moreover it is assumed that domestic dogs can be abandoned nd
consequently the become stray in the respective class. It is also assumed that
only domestic dogs reproduce themself and that the stray dogs of all class are
culled at the same rate. For the human population it is coherently supposed that
only exposed ones are treated and so they move in the vaccinated class. It can
be presented as:

\begingroup
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s4}
	\begin{cases}
		S_0'(t) = lS_1 - (\mu + c)S_0 - \beta_0S_0I_0 + p_0E_0 + \delta_1V_0\\
		E_0'(t) = lE_1 + \beta_0S_0I_0 - (\mu + c + \sigma_0 + p_0)E_0\\
		I_0'(t) = \sigma_0E_0 + I_1 - (\mu + c + \alpha)I_0\\
		V_0'(t) = lV_1 - (\mu + c + \delta_1)V_0\\
		S_1'(t) = A - \beta S_1I_0 - \beta_1S_1I_1 - (1 + \nu + l)S_1 + p_1E_1 + \delta2V_1\\
		E_1'(t) = \beta S_1I_0 + \beta_1S_1I_1 - (1 + \nu + \sigma_1 + l + p_1)E_1\\
		I_1'(t) = \sigma_1E_1 - (1 +  + k)I_1\\
		V_1'(t) = \nu(S_1 + E_1) - (1 + l + \delta_2)V_1\\
		S_h'(t) = H - \mu_hS_h - \lambda_{1h}S_hI_0 - \lambda_{2h}S_hI_1 + \delta_{1h}Eh + \delta_{2h}V_h\\
		E_h'(t) =\lambda_{1h}S_hI_0 + \lambda_{2h}S_hI_1 - (\mu_h + \sigma_h + \delta_{1h} + \nu_h)E_h\\
		I_h'(t) = \sigma_hE_h - (\mu_h + \alpha_h )I_h\\
		V_h'(t) = \nu_hE_h - (\mu_h + \delta_{2h})V_h\\
	\end{cases}
\end{align}
\endgroup
	
It can be proved that set 
\begin{align*}
  X=&\Bigl\{(S_0, E_0, I_0,V_0, S_1, E_1, I_1,V_1, S_h, E_h, I_h,V_h )\in\mathbb{R}_+^{12}~\vert\\
  &E_0, I_0, E_1, I_1, E_h, I_h,V_h\in\mathbb{R}_+\cup\{0\}\Bigr\}\\
\end{align*}
the region
\begin{align*}
\Gamma =&\Bigl\{p\in X~\vert~S_0 + E_0 + I_0 + V_0 + S_1 + E_1 + I_1 +\\
& V_1 + S_h + E_h +I_h+V_h\le\frac{A+H}{\min\{\mu + c, d,\mu_h\}}\Bigr\}\\
\end{align*}
is positively invariant.

It is proved that there are two equilibria 
\begin{itemize}
	\item a disease free equilibrium
	$E_0=(S_{0,0},0,0,V_{0,0},S_{1,0},0,0,$ $V_{1,0},S_{h,0},0,0,0)$ where the
	complete expression is reported in \cite{paperNostro};
	\item an endemic equilibrium $E*$ whose expression is  to complicate to
	formally compute, which can be proved to exist if $R_0>1$\cite{papernostro}.
\end{itemize}

After having computed $R_0$, whose expression is presented in
\cite{paperNostro}, it can be proved that if $R_0<1$, $E_0$ is is globally
asymptotically stable for the system \ref{s4}.

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
source("SEIR_dogs_stray_humans.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme +
  ylab(TeX("$I_h$")) + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
ggplot(data = I_h_df_long_1, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$l$, $\\epsilon$"), labels = labels_1) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = "", labels = c(
    "No intervention", "50% stray dog culling", "100% stray dog culling",
    "50% vaccination", "70% vaccination")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)") + guides(col = guide_legend(nrow = 3))
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
ggplot(data = I_h_df_long_3, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$\\nu_h$"), labels = nu_h_vect) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)")
```

## Seasonal

It is quite evident that the assumption of a constant recruitment rate it is not
biologically plausible \cite{54biopapnostro}. Moreover it has been observed that
also the contact rate and vaccination rate depends on time
\cite{78,biopapnostro}. Therefore there is a periodic component in the rabies
epidemic phenomenon that the previous model do not grasp. A new model has been
presented to highlight the periodicity component. It is a SEIR model with
periodic transmission rate:
\begingroup
\setlength{\mathindent}{0pt}
\begin{align}
	\label{s5}
	\begin{cases}
		S_d'(t) = A + \lambda_dR_d + \sigma_d (1 - \gamma_d )E_d - \beta_d(t) S_dI_d - (m_d + k_d)S_d\\
		E_d'(t) = \beta_d(t) S_dI_d - \sigma_d (1 - \gamma_d )E_d - \sigma_d\gamma_d E_d - (m_d + k_d)E_d\\
		I_d'(t) = \sigma_d\gamma_d E_d - (m_d + \mu_d)I_d\\
		R_d'(t) = k(S_d + E_d) - (m_d + \lambda_d)R_d\\
		S_h'(t) = B + \lambda_hR_h + \sigma_h(1 - \gamma_h )E_h - m_hS_h - \beta_{dh}(t)S_hI_h\\
		E_h'(t) = \beta_{dh}(t)S_hI_d - \sigma_h(1 - \gamma_h )E_h - \sigma_h\gamma_hE_h - (m_h + k_h )E_h\\
		I_h'(t) = \sigma_h\gamma_hE_h - (m_h + \mu_h)I_h\\\
		R_h'(t) = k_hE_h - (m_h + \lambda_h )R_h
	\end{cases}
\end{align}
\endgroup

where $\beta_d(t)=a[1 + bsin( \frac{\pi}{6} t + 5.5)]$ and
$\beta_{dh}(t)=a_1[1 + b_1sin( \frac{\pi}{6} t + 5.5)]$. Even in this case, set 
$$
X=\Bigl\{(S_d, E_d, I_d, R_d, S_h, E_h, I_h, R_h )\in (\mathbb{R}_+\cup\{0\})^8\Bigr\}
$$

it can be proved that there is a positively invariant region
\begin{align*}
\Gamma =&\Bigl\{p\in X~\vert~0 < S_d + E_d + I_d + R_d \le \frac{A}{m_d}\text{ and }\\
&0 < S_h + E_h + I_h + R_h \le \frac{B}{m_h}\Bigr\}\\
\end{align*}
It is proved that the system \ref{s5} has 
\begin{itemize}
	\item a disease free equilibrium $E_0=(S_{0},0,0,R_{0},S_{h},0,0,0)$ where the
	complete expression is reported in \cite{paperNostro}, which is globally
	asymptotically stable for $R_0<1$;
	\item at least one positive periodic solution \cite{111paperNostroBio}.
\end{itemize} 

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
source("SEIR_seasonal.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
  ylab(TeX("$I_h$")) + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
ggplot(data = as.data.frame(SEIR_out_2)) +
  geom_line(mapping = aes(time, I_h), color = "red") + myTheme + 
  ylab(TeX("$I_h$")) + xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
ggplot(data = I_h_df_long_1, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$S_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)")
```

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
ggplot(data = I_h_df_long_2, aes(x = time, y = value, colour = variable)) +
  geom_line() + scale_color_discrete(name = TeX("$I_d(0)$")) +
  myTheme + ylab(TeX("$I_h$")) + theme(legend.position="bottom") + 
  xlab("time (months)")
```


## Patch

MI: eq
MC: plot

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
source("SEIV_patch.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h_1), color = "blue") + 
  geom_line(mapping = aes(time, I_h_2), color = "red") + myTheme

```

# New stuff

MC : eq, plots, flow

## Patch + stray

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
source("SEIV_patch_stray.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h_1), color = "blue") + 
  geom_line(mapping = aes(time, I_h_2), color = "red") + myTheme

```

## Patch + stray + seasonal

```{r, echo=FALSE, fig.width=7, fig.height=6, fig.cap="\\label{fig:dsh1}plotting example"}
source("SEIV_patch_stray_seasonal.R")

ggplot(data = as.data.frame(SEIR_out)) +
  geom_line(mapping = aes(time, I_h_1), color = "blue") + 
  geom_line(mapping = aes(time, I_h_2), color = "red") + myTheme

```

# Conclusion

MI

\begin{thebibliography}{1}
	
	%Formato PLOS	
	
	\bibitem{Wang}Wang L, Tang Q, Liang G. Rabies and rabies virus in wildlife in mainland China, 1990-2013. Int J Infect Dis. 2014; 25:122-9.
	
	\bibitem{Schnell}Schnell MJ, McGettigan JP, Wirblich C, Papaneri A. The cell biology of rabies virus: using stealth to reach the brain. Nat Rev Microbiol. 2010; 8(1):51-61.
	
	\bibitem{Velasco}Velasco-Villa A, Escobar LE, Sanchez A, Shi M, Streicker DG, Gallardo-Romero NF, Vargas-Pino F, Gutierrez-Cedillo V, Damon I, Emerson G. Successful strategies implemented towards the elimination of canine rabies in the Western Hemisphere. Antiviral Res. 2017; 143:1-12.
	
	\bibitem{Hemachudha}Hemachudha T, Ugolini G, Wacharapluesadee S, Sungkarat W, Shuangshoti S, Laothamatas J. Human rabies: neuropathogenesis, diagnosis, and management. Lancet Neurol. 2013; 12(5):498-513.
	
\end{thebibliography}
